---
title: "Comparing social needs for transport, gaps in transit service and the Mobility Index for the west of Ireland"
runningheader: "The west of Ireland:transit needs-gaps" # only for pdf output
author: "Dr James Reynolds"
date: "`r Sys.Date()`"
bibliography: References.bib
csl: https://www.zotero.org/styles/chicago-fullnote-bibliography-16th-edition
suppress-bibliography: yes
output:
#  tufte::tufte_html: default
  tufte::tufte_handout:
#    citation_package: natbib
#    latex_engine: xelatex
#  tufte::tufte_book:
#    citation_package: natbib
#    latex_engine: xelatex

---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytransit)
library(sp)
library(strayr)
library(ptinpoly)
library(magrittr)
library(ggplot2)
library(sf)
library(ASGS.foyer)
library(raster)
library(ggmap)
library(units)
library(janitor)
library(mapview)
library(ggstatsplot)
library(gtsummary)
library(moments)
library(scales)
library(gtfstools)
library(lubridate)
library(kableExtra)
library(knitr)
library(readxl)
library(readr)
library(dplyr)
library(devtools)
library(gtfssupplyindex)
library(readabs)
library(gglorenz)
library(DescTools)
library(RColorBrewer)
library(lsr)
library(ggpubr)
library(viridis)
library(geosphere)
library(png)
library(absmapsdata)
library(tidytransit)
library(patchwork)
library(ggmapinset)
library(tufte)
library(gglorenz)
library(ggExtra)
library(ggthemes)
library(geomtextpath)
library(ggrepel)
#library(memisc, warn.conflicts = FALSE, quietly=TRUE)

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

```

# Introduction

# Research Background
Many
indexes, 
metrics 
and other indicators 
have been developed by 
researchers, 
practitioners 
and others 
to examine 
transport,
land use 
and related topics. 
Examples include: 
*Walk Score* and *Transit Score*[@WalkScore:2023tg], which rate a location out of 100 for walkability and 
transit accessibility respectively; 
metrics within the *Transit Capacity and Quality of Service Manual (TCQSM)*[@TCQSM:2013]; and
the UK's 
Public Transport Accessibility Level (PTAL) approach
and Public Transport Accessibility Index (PTAI)[@Saghapour:2016aa].

What may sometimes 
be lacking is 
comparison of 
the results from 
one set of indicators 
to another other.
In part this can be because of the 
variability in data availability. 
Censuses often provide a 
key resource for 
understanding transport-related issues, 
yet collection practices used, 
questions asked and 
other elements 
tend vary significantly from 
one jurisdiction to another. 
Likewise, 
transport agencies, 
government departments 
and other sources of data 
might have different 
approaches with respect 
to data sharing, 
open data policies 
or variability in 
what datasets they collect 
or produce 
in the first place. 
Hence, not all indicators 
can be calculated 
in all jurisdictions.  
That said,  
many indicators 
have been developed 
for specific contexts. 
An example is the 
recently reported
a Mobility Index (MI)[@McHenry:2023aa] 
for rural centres in the west of Ireland, 
developed because other MIs 
tend to be for more urbanised locations
where origins and destinations 
tend to be closer together.

Finally, 
some indicators are 
difficult or impossible 
to calculate independently. 
Some are closed-source 
and can only be obtained from a single entity^[e.g. WalkScore and Transit Score cannot be independently calculated and the exact algorithm remains unpublished].
Others may just be challenging to calculate 
due to a lack of, 
or free access to, 
software tools. 

While censuses 
and the collection of 
many other datasets 
at local, 
regional and 
national levels 
appear likely to remain 
specific to 
their own geographic contexts, 
there have been some recent 
developments in technology 
and transport 
practices that allow for
universal 
(or at least internationally-applicable) 
tools and approaches. 
Wong (2013) [@Wong:2013aa] 
developed software 
to calculate many of the *TCQSM* indicators from 
General Transit Feed Specification (GTFS) datasets, 
which is freely available online. 
This is possible because of the development 
and increased availability of GTFS feeds, 
which were originially introduced
to support the provision of transit 
directions on the Google Maps platform. 
This allows transit 
operators to publish 
their static timetable data 
in an open, 
simple 
and text-based format, 
which has been adopted by 
over 10,000 agencies[@GTFS] 
Hence, software tools 
that take GTFS data 
as inputs 
can generally be adapted 
easily and rapidly 
to many systems and 
locations, 
as in how Wong (2013) 
compared TCQSM metrics across 50 USA transit operators. 

## Social needs-gap approach

Another such recently developed 
software tool is the 
*gtfssupplyindex* 
R package [@Reynolds2025gtfsmainpaper],
which builds on 
other GTFS-related R packages 
including 
*gtfstools* [@R-gtfstools]; 
and *tidytransit* [@R-tidytransit]. 
The *gtfssupplyindex* can 
be used to calculate a 
transit Supply Index (SI), 
developed
in research related 
to identifying 
gaps in public transport services 
on the basis of social needs for transport reported in Currie (2010)
[@currie2010identifying]. 
The SI equation is:

$$SI_{area, time} = \sum_{i=1}^{n}{\frac{Area_{Bi}}{Area_{area}}SL_{i, time}}$$

where:

-   $SI_{area, time}$ is the Supply Index for the *area of interest* and a given period of time;

-   $Area_{Bi}$ is the amount of the buffer zone for each stop (i) that is within the *area of interest*^[In @currie2010identifying this buffer zone was based on a radius of 400 metres for bus and tram stops, and 800 metres for railway stations. The same definition is used here.];

-   $Area_{area}$ is the area of each *area of interest*; and

-   $SL_{i,time}$ is the number of transit arrivals for each stop (i) within the given time period.

Currie (2010) also used 
a Composite Social Needs Index, 
based mostly on census data available from the Australian Bureau of Statistics (ABS), including measures 
related to income, age, car ownership and other 
indicators related to the extent to which 
those living in each area of interest might 
have higher or lower needs for public transport^[These included the Index of Relative Socio-Economic Advantage/Disadvantage (IRSAD), 
which is reported directly 
by the ABS and
based on 31 weighted indicators 
and data collected during the census. A transport-specific need index, 
was also included, based on eight weighted indicators: adults without cars (0.19), 
distance to the Melbourne Central Business District (0.15), 
persons aged over 60 years (0.14), 
persons on a disability pension (0.12), 
low income households (0.10), 
adults not in the labour force (0.09), 
students (0.09) and 
persons 5-9 years (0.12).
The IRSAD and transport-specific needs index were then weighted by population and combined to form the Social Needs Index.]

The final step 
of the Currie (2010) 
social needs-gap approach 
involved identifying those 
areas of interest that had 
both high social needs, 
but low or zero transit supply levels. 
Hence, the result is 
a comparative assessment amongst all of the areas of interest
that allows those that would have the largest gaps 
between the social need for transport to be identified, 
rather than being based on comparing raw scores to a threshold.  
Its potential application 
to policy-making, 
route design 
and service planning 
was suggested to be 
“substantially more useful than the presentation of anecdotal evidence, which is the most common means of identifying transport needs in local transport studies throughout the world”[@currie2010identifying]. 
Essentially, use of this approach might help practitioners 
identify where best the available, or any additional, resources might be allocated^[As opposed to planning routes based on minimum thresholds, such as a maximum allowable distance between residential properties and a transit stop, or frequency standards based on ridership or population.]

A feature of the SI equation is that scores can be summed across areas of interest so as to provide an aggregate score 
for a larger district, region or city.
The SI also had the advantage of including 
the frequency 
and coverage 
of transit services, 
meaning that the score 
reflects the accessibility 
'to' public transport for those living within an area, 
but also how accessible the area is 'by' transit.

The development of the Currie (2010) needs-gap approach
predated the GTFS, 
and hence the analysis required 
bespoke cleaning 
and processing 
of data obtained directly 
from the transit agency. 
@Reynolds2025gtfsmainpaper,
however, reported the development of 
the *gtfssupplyindex*, 
which allows SI scores to be 
calculated directly from GTFS files, 
and results from applying this to 
Melbourne, Australia. 
Beyond this, 
however,
the *gtfssupplyindex* 
package has not yet been 
applied to other locations, 
or indeed rural geographies.
Also unclear is how 
such results might compare 
to other transport-related indicators, 
including the Mobility Index for rural areas 
developed by McHenry et al (2023)[@McHenry:2023aa]


In the following sections, 
therefore, 
this paper 
reports the application 
of the *gtfssupplyindex* 
package to the west of Ireland, 
adopting the same 
case and context as McHenry et al (2023). 
Results from a social needs-gap 
analysis are compared to 
those of McHenry et al (2023) 
so as to provide insights 
into the differences 
between these indicators. 
This research is 
anticipated to be 
part of a larger 
programme that 
would continue to 
apply different indicators and approaches 
to this case, 
so as to allow 
deeper understandings 
of both transport in the west of Ireland, 
but also of how well 
existing indicators and other analysis techniques 
might allow 
researchers, 
practitioners and policy-makers 
to understand and 
improve rural transport networks.  

# Methodology
The *areas of interest* adopted 
in this study 
are the 
2022 definition of *Small Areas*, 
as used in the 
Irish Census^[However, 
data from the Irish Census 
is reported seperately 
for Galway City and the rest of County Galway("Galway City" is included as an additional county in the COUNTY_ENGLISH field), 
whereas in this analysis 
Small Areas within Galway City 
are reported as part of County Galway.]
GTFS data for this paper was sourced from
@Transport-for-Ireland:0aa on October 1, 2025^[Selecting the "All GTFS Operators" feed]. 
The time period selected was the seven days starting October 1, 2025.

The same area of interest categorisation approach as in 
Currie (2010) 
and Reynolds et al (2025) 
was adopted^[Transit supply is grouped into seven categories, based on first allocating all areas of interest with a SI score of zero into "Zero Supply". The average SI score of the remainder is then used to split the rest of the areas of interest, which are then further subdivided into three even groups ("Very Low", "Low", "Below average") below and three even groups above ("Above average", "High", "Very High) the mean. A similar categorisation is used to convert the Composite Needs Index scores into six, rather than seven categories (as there are no 'Zero' scores for need).]



```{r load_ireland_gtfs, eval = FALSE, echo = FALSE}

## Check feed loads as tidytransit
#TFI_GTFS_All_251001 <- tidytransit::read_gtfs("data/251001/TFI_GTFS_All_251001.zip")
#TFI_GTFS_All_251001 %>% class()
## Yes feed loads successfully

TFI_GTFS_All_251001_list_gtfs <- gtfssupplyindex::gtfs_by_route_type("data/251001/TFI_GTFS_All_251001.zip")

```

```{r load_region_of_interest_function, eval = TRUE, echo = FALSE}

load_region_of_interest <- function(){
region_of_interest <- Small_Area_National_Statistical_Boundaries_2022 %>% 
  filter(COUNTY_ENGLISH == "CLARE" |
           COUNTY_ENGLISH == "DONEGAL" |
           COUNTY_ENGLISH == "GALWAY" |
           COUNTY_ENGLISH == "LEITRIM" |
           COUNTY_ENGLISH == "MAYO" |
           COUNTY_ENGLISH == "ROSCOMMON" |
           COUNTY_ENGLISH == "SLIGO" | 
           COUNTY_ENGLISH == "GALWAY CITY")
return(region_of_interest)}
```

```{r run_for_all_modes_ireland_, eval = FALSE, echo = FALSE}

list_gtfs = TFI_GTFS_All_251001_list_gtfs


Small_Area_National_Statistical_Boundaries_2022 <- st_read("data/Small_Area_National_Statistical_Boundaries_2022_Ungeneralised_view_-7354763930310470674.gpkg")
st_geometry(Small_Area_National_Statistical_Boundaries_2022) <- "SHAPE"

region_of_interest <- load_region_of_interest()

region_of_interest <- st_set_geometry(region_of_interest, "SHAPE")


#Load region_of_interest codes
areas_of_interest <- load_areas_of_interest(region_of_interest,  
  area_id_field = "SA_PUB2022")

areas_of_interest <- st_set_geometry(areas_of_interest, "geometry")


stops_in_or_near_areas <- gtfssupplyindex:::stops_in_walk_dist(
  list_gtfs = list_gtfs, 
  areas_of_interest = areas_of_interest,
  EPSG_for_transform = 2157,
  verbose = FALSE
)



si_west_ireland_week_251001_by_day_hour_route <- SI_by_day_hour_and_route_type(
     list_gtfs = list_gtfs, 
stops_in_or_near_areas = stops_in_or_near_areas,
 start_date_ymd = "2025-10-01", 
 end_date_ymd = "2025-10-07",
verbose = TRUE)

write.csv(si_west_ireland_week_251001_by_day_hour_route, "results/si_west_ireland_week_251001_by_day_hour_route.csv")


si_west_ireland_week_251001_aggregated <- si_west_ireland_week_251001_by_day_hour_route %>% 
           group_by(area_id) %>%    
           summarize(SI = sum(SI))

write.csv(si_west_ireland_week_251001_aggregated, "results/si_west_ireland_week_251001_aggregated.csv")


```


For assessing social needs for transport, it is not possible to exactly replicate the Composite Social Needs Index used in the Currie 2010 or Reynolds et al (2025) analyses. These used indicators drawn from Australian censuses, not all of which have equivalents in the Irish censuses or other available datasets. 
Therefore, assessment of social needs for transport for Small Areas within the west of Ireland is based on various indicators reported in the 2022 Irish Census^[Item T15_1_NC is the number of households without a motor car. Items T1_1_AGE60-64T, T1_1_AGE65-69T, ... T1_1_AGE80-84T and T1_1_AGE85T report persons aged 60 years and older; T8_1_UTWSDT reports the total number of people unable to work due to permanent sickness or disability; T8_1_TT reports the total population aged 15 years and over, while T8_1_WT reports the total at work, thereby allowing calculation of the number of adults (15+) not in the labour force; T8_1_ST reports the number of people aged 15 years and over who are students; and T1_1AGE5T, T1_1AGE6T ... T1_1AGE9T report the number of people aged 5 through 9. There does not appear to be any data available about low income households at the "Small Area" geographic area. As well, there does not appear to be a IRSAD or similar socio-economic indicator available.]. As such, this analysis adapts the Currie (2010) approach and uses the same weighting for the following indicators: households without cars (0.19); distance to the centre of the nearest urban area with population of at least 4,500 people^[As the population of Carrick-On-Shannon is a little below 5,000 this threshold ensures that there is at least one 'centre' in each County] (0.15); persons aged 60 years and over (0.14), persons unable to work due to permant sickness or disability (0.12), those aged 15 years or older who are not at work; students (0.09) and people aged 5-9 years (0.12).  Each of these indicators were normalised across all of the Small Areas in the west of Ireland, weighted as indicated to develop an Social Index score. As in Currie (2010), the Social Index score was then normalised, weighted by the total population in each Small Area and normalised to produce a Relative Social Index score. The (unweighted) Social Index score and the Relatvie Social Index score were then added together to produce the Composite Social Needs Index.


```{r read_census_data_2021, fig.show="hold", eval= TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=TRUE, fig.fullwidth = TRUE, fig.cap="SI scores by SA3, census day 2016 and 2021"}

SAPS_2022_Small_Area_UR_171024 <- read_csv("data/SAPS_2022_Small_Area_UR_171024.csv")

SAPS_2022_Small_Area_UR_171024 <- SAPS_2022_Small_Area_UR_171024 %>% 
  clean_names()

social_needs <- SAPS_2022_Small_Area_UR_171024 %>% 
  select(guid, geogid, geogdesc, ur_category, ur_category_desc) 
 
 
social_needs$over_60 <- SAPS_2022_Small_Area_UR_171024 %>% 
  select(t1_1age60_64t, t1_1age65_69t, t1_1age70_74t, t1_1age75_79t, t1_1age80_84t, t1_1agege_85t) %>% 
  rowSums()


social_needs$age_five_to_nine <- SAPS_2022_Small_Area_UR_171024 %>% select(
  t1_1age5t, t1_1age6t, t1_1age7t, t1_1age8t, t1_1age9t) %>%
  rowSums()

social_needs$adult_no_car <- NA
social_needs$low_income_families <- NA

social_needs$adult_not_labour_force <- SAPS_2022_Small_Area_UR_171024$t8_1_tt - 
  SAPS_2022_Small_Area_UR_171024$t8_1_wt

social_needs$students <- SAPS_2022_Small_Area_UR_171024$t8_1_st

social_needs$IRSAD <- NA

social_needs$population <- SAPS_2022_Small_Area_UR_171024$t1_1agett


##Rescaling to 0-100 and adding the accessibility indicator needs to be done on a areas_of_interest by areas_of_interest bias, as which areas are included will matter. 

```





```{r social_needs_calc, fig.dim=c(7.5,7.5), echo = FALSE, eval=TRUE, warning=FALSE, message=FALSE, cache=FALSE,  fig.fullwidth = FALSE,fig.cap="Distribution of categories of social need index scores"}


#Load region_of_interest codes
Small_Area_National_Statistical_Boundaries_2022 <- st_read("data/Small_Area_National_Statistical_Boundaries_2022_Ungeneralised_view_-7354763930310470674.gpkg")
st_geometry(Small_Area_National_Statistical_Boundaries_2022) <- "SHAPE"

#run load_region_of_interest function

region_of_interest <- load_region_of_interest()

#Correct Galway City Small Areas to be in County Galway in the County_English field
region_of_interest$COUNTY_ENGLISH <- if_else(region_of_interest$COUNTY_ENGLISH == "GALWAY CITY", "GALWAY", region_of_interest$COUNTY_ENGLISH)

##Change NA SA_URBAN_AREA_NAME to "non-urban"
#region_of_interest$SA_URBAN_AREA_NAME <- region_of_interest$SA_URBAN_AREA_NAME %>% 
#  replace_na("non-urban")

region_of_interest <- region_of_interest %>%
  select(SA_GUID_2022, SA_PUB2022, SA_URBAN_AREA_NAME, SA_NUTS2_NAME, SA_NUTS3_NAME, COUNTY_ENGLISH)

names(region_of_interest) <- c("SA_GUID_2022", "SA_PUB2022", "SA_URBAN_AREA_NAME", "SA_NUTS2_NAME", "SA_NUTS3_NAME", "COUNTY_ENGLISH", "geometry" )
region_of_interest <- region_of_interest %>% st_set_geometry("geometry")

#Load region_of_interest codes
areas_of_interest <- load_areas_of_interest(region_of_interest,  
  area_id_field = "SA_PUB2022")


### Calculate distances to the post offices
centroids <- areas_of_interest %>% 
  st_centroid()

# si_pop_by_urban_area %>% filter(population >= 4500) provides list of places. Select central post office in each as 'centre' of each Urban area

#Athenry \"Athlone\", \"Ballina\", \"Ballinasloe\", \"Ballybofey-Stranorlar\", \"Buncrana\", \"Carrick-On-Shannon\", \"Carrick-On-Shannon\", \"Castlebar\", \"Ennis\", \"Galway city and suburbs\", \"Letterkenny\", \"Loughrea\", \"Oranmore\", \"Roscommon\", \"Shannon\", \"Sligo\", \"Tuam\", \"Westport\")"

post_offices <- data.frame(
  centre = c("Athenry", "Athlone", "Ballina", "Ballinasloe", "Ballybofey-Stranorlar", "Buncrana", "Carrick-On-Shannon", "Castlebar", "Ennis", "Galway city and suburbs", "Letterkenny", "Loughrea", "Oranmore", "Roscommon", "Shannon", "Sligo", "Tuam", "Westport"),
  lat = c(53.29922090590193, 53.42383747922755, 54.11282845042087, 53.329553855273986, 54.79974773020907, 55.13519888166909, 53.94550169284656, 53.855000000857665, 52.8455723784013, 53.27512525358046, 54.9512675471456,  53.1991052895609, 53.273602789495875, 53.63011287294085, 52.710128301270466, 54.27274270646182, 53.51562650727329, 53.80111936724792 ),
  lon = c(-8.745304658788108, -7.94397471189948, -9.156154674069866, -8.221171062711592, -7.7787459655676, -7.454363243432908, -8.096176594681543, -9.302476398672507,  -8.985093169375743, -9.051859277468566, -7.73591409935381, -8.568511261067886, -8.929322726854224, -8.18972171613282, -8.878905148835077, -8.47547944417072, -8.851331129710097, -9.520243407006244)
)

post_offices <- post_offices %>% st_as_sf(coords = c("lon", "lat"), crs = 4326)
post_offices <- st_transform(post_offices, crs=2157)

post_office_distances <- vector()
i <- 1
for(i in 1:nrow(centroids)){
  post_office_distances[i] <- centroids[i,] %>% 
    st_distance(post_offices) %>% as.numeric() %>% min()
  }

region_of_interest$distance_m_to_GPO <- post_office_distances  

social_needs <- region_of_interest %>% left_join(social_needs %>% st_drop_geometry(), by = join_by(SA_GUID_2022 == guid))


#ggplot() + 
#  geom_sf(data = areas_of_interest, aes(fill=dist_to_nearest_major_centre)) +
#  geom_sf(data = post_offices, aes(label=centre))


#FUNCTION TO CALCULATE COMBINED NEEDS INDEX
#Comment out indicators we don't have
calculate_IRSAD_and_need_index <- function(social_dataset) {
 # scale all to 0 to 100
 # social_dataset$IRSAD_100 <- rescale_max(
#     social_dataset$IRSAD, to = c(0, 100))
#  social_dataset$adult_no_car_100 <- rescale_max(
#    social_dataset$adult_no_car, to = c(0, 100))
   social_dataset$accessibility_100 <- rescale_max(
     social_dataset$distance_m_to_GPO, to = c(0, 100))
   social_dataset$over_60_100 <- rescale_max(
     social_dataset$over_60, to = c(0, 100))
#   social_dataset$disability_pension_100 <- rescale_max(
#     social_dataset$disability_support_pension, to = c(0, 100))
#   social_dataset$low_income_families_100 <- rescale_max(
#     social_dataset$low_income_families, to = c(0, 100))
  social_dataset$adult_not_labour_force_100 <- rescale_max(
     social_dataset$adult_not_labour_force, to = c(0, 100))
  social_dataset$students_100 <- rescale_max(
     social_dataset$students, to = c(0, 100))
  social_dataset$age_five_to_nine_100 <- rescale_max(
     social_dataset$age_five_to_nine, to = c(0, 100))
  
  # calculate needs scores and scale to 100
  social_dataset$needs_score <- social_dataset$accessibility_100 * 0.15 + 
    social_dataset$over_60_100 * 0.14 + 
    social_dataset$adult_not_labour_force_100 * 0.09 + 
    social_dataset$students_100 * 0.09 + 
    social_dataset$age_five_to_nine_100 * 0.12
  
  social_dataset$needs_score_100 <- rescale_max(
     social_dataset$needs_score, to = c(0, 100))

  
  
  ### No IRSAD, so will only return the needs score based on the available transport indicators 
# weighted by population 
  
social_dataset$needs_score_100_population_weighted <- rescale_max(social_dataset$needs_score_100  *  social_dataset$population, to = c(0,100))
  

#  social_dataset$total_need_IRSAD <- social_dataset$IRSAD *
#    social_dataset$population
#  social_dataset$total_transport_need <- social_dataset$needs_score * 
#    social_dataset$population
 
#  # create combined indicator 0 to 100
#  social_dataset$combined_needs_index <- social_dataset$total_need_IRSAD + 
#    social_dataset$total_transport_need 
#  social_dataset$combined_needs_index <- rescale_max(
#     social_dataset$combined_needs_index, to = c(0, 100))
  
social_dataset$combined_needs_index <- rescale_max(
     social_dataset$needs_score_100_population_weighted + social_dataset$needs_score_100, to = c(0, 100))

  
  return(social_dataset %>% 
           select(area_id, combined_needs_index, population)) 
}

social_needs <- social_needs %>% select(SA_PUB2022, distance_m_to_GPO, over_60, age_five_to_nine, adult_no_car, low_income_families, adult_not_labour_force, students, IRSAD, population, geometry)

names(social_needs) <- c("area_id", "distance_m_to_GPO", "over_60", "age_five_to_nine", "adult_no_car", "low_income_families", "adult_not_labour_force", "students", "IRSAD", "population", "geometry")

needs_index <- calculate_IRSAD_and_need_index(social_needs)


## FUNCTION to define thresholds and return Very High, High etc. 
# as per Currie2010 Table 3 and page 34. 
# the approach appears to be to first identify zones that have zero supply 
# the remainder are then split into those above and below the average SI value
# these two groups are then each split into three groups of roughly equal size. 
set_thresholds <-function(score_by_area_dataframe){ 

names(score_by_area_dataframe) <- c("area_id", "score")  
  
## Define Very High, High etc thresholds
cuts_lower_groups <- score_by_area_dataframe %>% 
  st_drop_geometry() %>% 
  filter(score !=0) %>%
  filter(score <= score %>% mean()) %>% 
  select(score) %>% unlist() %>% 
  as.vector() %>% 
  quantileCut(3)

cuts_upper_groups <- score_by_area_dataframe %>% 
  st_drop_geometry() %>% 
  filter(score !=0) %>%
  filter(score > score %>% mean()) %>%
  select(score) %>% unlist() %>% 
  as.vector() %>% 
  quantileCut(3)

# Recode lower_group factors
allocation_lower_groups <- cuts_lower_groups %>% fct_recode(
  "Very Low" = levels(cuts_lower_groups)[1], 
  "Low" = levels(cuts_lower_groups)[2], 
  "Below average" = levels(cuts_lower_groups)[3])

# Recode upper_group factors
allocation_upper_groups <- cuts_upper_groups %>% fct_recode(
  "Above average" = levels(cuts_upper_groups)[1], 
  "High" = levels(cuts_upper_groups)[2], 
  "Very High" = levels(cuts_upper_groups)[3])


# Join back to lower_group SIs and areas_of_interest
score_by_area_dataframe_below_average <- score_by_area_dataframe %>% 
  st_drop_geometry() %>% 
  filter(score !=0) %>%
  filter(score <= score %>% mean()) 
score_by_area_dataframe_below_average$transit_supply <- allocation_lower_groups

# Join back to upper_group SIs and areas_of_interest
score_by_area_dataframe_above_average <- score_by_area_dataframe %>% 
  st_drop_geometry() %>% 
  filter(score !=0) %>%
  filter(score > score %>% mean()) 
score_by_area_dataframe_above_average$transit_supply <- allocation_upper_groups

# combine upper and lower groups
  
score_by_area_dataframe_non_zero <- add_row(score_by_area_dataframe_below_average, 
                                         score_by_area_dataframe_above_average)

#join back to dataframe, and put "Zero" in N/A
score_by_area_dataframe <- full_join(
  score_by_area_dataframe %>% 
    st_drop_geometry(), 
  score_by_area_dataframe_non_zero
)

score_by_area_dataframe$transit_supply <- score_by_area_dataframe$transit_supply %>% 
  fct_explicit_na(na_level = "Zero Supply") %>% 
  fct_shift(n = -1)

#define thresholds as numeric using https://stackoverflow.com/questions/39387745/cut-extracting-the-minimum-and-maximum-of-cut-levels-as-columns-in-data-frame
thresholds <- tibble(cuts_lower_groups %>% as_tibble()  %>% 
  mutate(x_tmp = str_sub(value, 2, -2)) %>% 
  separate(x_tmp, c("min", "max"), sep = ",") %>% 
  mutate_at(c("min", "max"), as.double) %>% 
  select(max) %>% unique() %>%
  arrange(max)
)

thresholds <- rbind(thresholds, cuts_upper_groups %>% as_tibble()  %>% 
  mutate(x_tmp = str_sub(value, 2, -2)) %>% 
  separate(x_tmp, c("min", "max"), sep = ",") %>% 
  mutate_at(c("min", "max"), as.double) %>% 
  select(max) %>% unique() %>%
  arrange(max)
)

#return dataframe and information about level cut offs
score_by_area_dataframe_and_cuts_dataframe <- list(
  si_by_area = score_by_area_dataframe,
  thresholds_text = c(levels(cuts_lower_groups), levels(cuts_upper_groups)), 
  thresholds_values = thresholds
  )
return(score_by_area_dataframe_and_cuts_dataframe)
}
# END FUNCTION

thresholds_holding <- needs_index %>%
  st_drop_geometry() %>% 
  select(area_id, combined_needs_index) 

thresholds_holding$combined_needs_index <- thresholds_holding$combined_needs_index %>% unlist()

thresholds_holding <- set_thresholds(thresholds_holding)
names(thresholds_holding[[1]]) <- c("area_id", "composite_needs_index", "composite_needs_category")
thresholds_holding[[1]]$composite_needs <- thresholds_holding[[1]]$composite_needs %>% fct_other(keep = c("Very Low", "Low", "Below average", "Above average", "High", "Very High"), other_level = "NA")

#join to guid geometry data
needs_index <- left_join(
  areas_of_interest %>%
    select(area_id),
  thresholds_holding[[1]]
)



combined_needs_2021_plot <- ggplot()+ 
  geom_sf(data=needs_index %>% na.omit(),
          aes(fill =  composite_needs_category), colour=NA) +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(), #remove y axis ticks
        legend.position="none"
  ) +
    #    legend.text = element_text(size = 4), 
    #    legend.title = element_text(size = 5)) + 
  scale_fill_manual(values = c("#FFFB07", "#CEE102", "#A3CF07", "#61AE09", "#2EA105", "#0A7B01")) +
    ggspatial::annotation_scale(location = 'tl', height = (unit(0.05, "cm")), text_cex = 0.25) 
#combined_needs_2021_plot

```


# Results


```{r threshold_function, fig.show="hold", out.width="100%", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE,  cache=FALSE, results='hide'}


## FUNCTION to set thresholds and return tibble, plot output and thresholds
si_thresholds_and_plot_function <- function(si_by_area, areas_of_interest){
# Set column names
names(si_by_area) <- c("area_id", "SI")
names(areas_of_interest) <- c("area_id", "geometry")
  
si_by_area$area_id <- si_by_area$area_id %>% 
  as.character()


##Join to areas_of_interest so as to identify (and enumerate as zero) 
# those areas of interest that have no SI score 

si_by_area <- full_join(
  areas_of_interest,
  si_by_area)
si_by_area[is.na(si_by_area)] <- 0

si_by_area_thresholds <- set_thresholds(si_by_area_dataframe = si_by_area)
si_by_area <- si_by_area_thresholds[[1]]


si_by_area <- left_join(
  areas_of_interest, 
  si_by_area)

output_plot <- ggplot()+ 
  geom_sf(data=si_by_area %>% na.omit(),
          aes(fill = transit_supply), colour=NA) +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(), #remove y axis ticks
        legend.position="none",
#        legend.text = element_text(size = 4), 
#        legend.title = element_text(size = 5), 
#        legend.key.size = unit(0.08, 'cm')
        ) + 
  scale_fill_manual(values = c("white", "#F9D8B1", "#F7C387", "#F29C33", "lightgreen", "#53A212", "darkgreen")) 


  outputs <- list("si_by_area" = si_by_area, 
                  "output_plot" = output_plot, 
                  "thresholds" = si_by_area_thresholds[[2]], 
                  "thresholds_values" = si_by_area_thresholds[[3]])
  return(outputs)
}

```

```{r region_of_interest_2025, fig.show="hold", out.width="100%", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE,  cache=FALSE, results='hide'}

### Load 2025 results
si_by_area <- read.csv("results/si_west_ireland_week_251001_aggregated.csv") %>% 
  as_tibble()
si_by_area <- si_by_area[,2:3]

si_region_of_interest_aggregated <- si_thresholds_and_plot_function(si_by_area, areas_of_interest)

## Connect SI results to area definitions
si_region_of_interest_aggregated$si_by_area <- left_join(
  si_region_of_interest_aggregated$si_by_area,
  region_of_interest %>% st_drop_geometry(), 
  by=join_by("area_id" == "SA_PUB2022"))


## Import population data
## T1_1AGETT is total population field.  
# GUID is identification field
SAPS_2022_Small_Area_UR_171024 <- read_csv("data/SAPS_2022_Small_Area_UR_171024.csv") %>% select(GUID, T1_1AGETT) 

names(SAPS_2022_Small_Area_UR_171024) <- c("GUID", "population")

si_region_of_interest_aggregated$si_by_area <- left_join(
  si_region_of_interest_aggregated$si_by_area,
  SAPS_2022_Small_Area_UR_171024, 
  by=join_by("SA_GUID_2022" == "GUID"))

# Categorise population into Very High, High groups etc.
region_of_interest_population <- si_region_of_interest_aggregated$si_by_area %>% 
  st_drop_geometry() %>% 
  select(area_id, population)
names(region_of_interest_population) <- c("area_id", "SI")
region_of_interest_population <- set_thresholds(region_of_interest_population)
names(region_of_interest_population[[1]]) <- c("area_id", "population", "population_category")

si_region_of_interest_aggregated$si_by_area$population_category <- 
  region_of_interest_population[[1]]$population_category

si_region_of_interest_aggregated$si_by_area$population_category <- 
  si_region_of_interest_aggregated$si_by_area$population_category %>%
  factor(levels=c("Very Low", "Low", "Below average", "Above average", "High", "Very High"))

#Calculate population density
si_region_of_interest_aggregated$si_by_area$population_density <- 
  si_region_of_interest_aggregated$si_by_area$population %>% st_drop_geometry() /
  si_region_of_interest_aggregated$si_by_area %>% st_area()
#Categorise population density  
region_of_interest_population_density <- si_region_of_interest_aggregated$si_by_area %>% 
  st_drop_geometry() %>% 
  select(area_id, population_density)
names(region_of_interest_population_density) <- c("area_id", "SI")
region_of_interest_population_density$SI <- region_of_interest_population_density$SI %>%
  as.numeric()
  
region_of_interest_population_density <- set_thresholds(region_of_interest_population_density)
names(region_of_interest_population_density[[1]]) <- c("area_id", "population_density", "density_category")

si_region_of_interest_aggregated$si_by_area$density_category <- 
  region_of_interest_population_density[[1]]$density_category

si_region_of_interest_aggregated$si_by_area$density_category <- 
  si_region_of_interest_aggregated$si_by_area$density_category %>%
  factor(levels=c("Very Low", "Low", "Below average", "Above average", "High", "Very High"))


si_region_of_interest_aggregated$si_by_area$population_density <-  si_region_of_interest_aggregated$si_by_area$population_density %>% as.numeric()

```


## SI scores and transit supply categories

Table 1 shows summary statistics for SI scores, and by transit supply category. 

```{marginfigure}
Table 1: Summary statistics; SI score by transit supply category, by Small Area.
```

```{r summary_table_west_ireland, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = FALSE}

theme_gtsummary_compact()
si_region_of_interest_aggregated$si_by_area %>% 
  st_drop_geometry() %>% 
  select(SI, transit_supply) %>%
   tbl_summary(
    by = transit_supply,
    digits = SI ~ 1,
    type = all_continuous() ~ "continuous2",
     statistic = list(all_continuous() ~ c("{min}", "{p25}", "{mean}", "{p75}", "{max}", "{sum}")
   )) %>%
    add_overall() %>% 
  modify_spanning_header(c("stat_1", "stat_2", "stat_3", "stat_4", "stat_5", "stat_6", "stat_7") ~ "**Transit supply category**") %>%
  as_kable_extra()

```



The lower and upper SI score thresholds for each of the transit supply categories are shown in Table 2, along with the number of Small Areas in each category. 

```{marginfigure}
Table 2: Transit supply category groupings, SI score thresholds and number of Small Areas in each category.
```


```{r summary_table_thresholds, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = FALSE}

thresholds_table <- tibble(Category = c("Zero Supply", "Very Low", "Low", "Below average", "Above average", "High", "Very High"), 
                           Lower = c(0, 0, si_region_of_interest_aggregated$thresholds_values[1,1] %>% unlist() %>% as.numeric(), si_region_of_interest_aggregated$thresholds_values[2,1] %>% unlist() %>% as.numeric(), si_region_of_interest_aggregated$thresholds_values[3,1] %>% unlist() %>% as.numeric(), si_region_of_interest_aggregated$thresholds_values[4,1] %>% unlist() %>% as.numeric(), si_region_of_interest_aggregated$thresholds_values[5,1] %>% unlist() %>% as.numeric()),
                           Upper = c(0, si_region_of_interest_aggregated$thresholds_values[1,1] %>% unlist() %>% as.numeric(), si_region_of_interest_aggregated$thresholds_values[2,1] %>% unlist() %>% as.numeric(), si_region_of_interest_aggregated$thresholds_values[3,1] %>% unlist() %>% as.numeric(), si_region_of_interest_aggregated$thresholds_values[4,1] %>% unlist() %>% as.numeric(), si_region_of_interest_aggregated$thresholds_values[5,1] %>% unlist() %>% as.numeric(), si_region_of_interest_aggregated$thresholds_values[6,1] %>% unlist() %>% as.numeric()), 
                           Areas = c(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SI == 0) %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Very Low") %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Low") %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Below average") %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Above average") %>% nrow(),si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "High") %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Very High") %>% nrow())
                          )

thresholds_table %>% kable(align = "lrrr", digits = c(0, 0, 0, 0), format.args = list(big.mark = ",", scientific = FALSE))
```


Some `r label_number(big.mark=",")(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SI == 0) %>% nrow())` of the 
`r label_number(big.mark=",")(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% nrow())` Small Areas in the west of Ireland (`r label_percent()(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SI == 0) %>% nrow() / si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% nrow())`) 
have SI scores of zero, implying that there is no transit service 
at all. Of those Small Areas that had at least some transit, `r label_number(big.mark=",")(sum(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Very Low") %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Low") %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Below average") %>% nrow()))`
(`r label_percent(accuracy=1)(sum(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Very Low") %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Low") %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Below average") %>% nrow())/si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% nrow())` of the total) had SI scores below the average 
score (`r label_number(big.mark=",")(si_region_of_interest_aggregated$thresholds_values[3,1] %>% unlist() %>% as.numeric())`), with the remaining 
`r label_number(big.mark=",")(sum(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Above average") %>% nrow(),si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "High") %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Very High") %>% nrow()))` (`r label_percent(accuracy=1)(sum(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Above average") %>% nrow(),si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "High") %>% nrow(), si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(transit_supply == "Very High") %>% nrow())/si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% nrow())` of the total Small Areas in the west of Ireland. 

A map and the distribution of transit Supply Index (SI) categories across the west of Ireland is shown in Figure \ref{fig:SI_map_region_of_interest_county}. This Figure also shows the population living within Small Areas in each category, including by County. 


```{r SI_map_region_of_interest_county_population, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across west of Ireland, by population"}


#Set boundings box
bbox_list <- region_of_interest %>% 
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()

pcol_si_region_of_interest <- si_region_of_interest_aggregated$si_by_area %>%
  st_drop_geometry() %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, population, fill = transit_supply), colour="black", size=0.2) +
  scale_y_continuous(limits = c(0,500000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  population+80000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 0.1)(perc)})")), size = 0.75) +
   coord_flip() + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "All west of Ireland (by pop.)",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.5),
    axis.text.y = element_text(size = 1.5, hjust = 1),
    axis.text.x = element_blank()
  ) 

pcol_si_region_of_interest_blank <- si_region_of_interest_aggregated$si_by_area %>%
  st_drop_geometry() %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, population, fill = transit_supply), colour="black", size=0.2) +
  scale_y_continuous(limits = c(0,500000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  population+80000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
   coord_flip() + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "All west of Ireland (by pop.)",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 




```


```{r SI_map_region_of_interest_county_clare, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across west of Ireland, by population"}


#Set boundings box
bbox_list <- region_of_interest %>% 
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()

pcol_si_county_clare <- si_region_of_interest_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(COUNTY_ENGLISH == "CLARE") %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/pcol_si_region_of_interest$data[1:7,2] %>% sum()) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, population, fill = transit_supply), colour="black", size=0.2) +
  scale_y_continuous(limits = c(0,500000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  population+80000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 0.1)(perc)})")), size = 0.75) +
   coord_flip() + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "County Clare (by pop.)",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.0),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 

```


```{r SI_map_region_of_interest_county_galway, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across west of Ireland, by population"}


#Set boundings box
bbox_list <- region_of_interest %>% 
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()

pcol_si_county_galway <- si_region_of_interest_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(COUNTY_ENGLISH == "GALWAY") %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/pcol_si_region_of_interest$data[1:7,2] %>% sum()) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, population, fill = transit_supply), colour="black", size=0.2) +
  scale_y_continuous(limits = c(0,550000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  population+80000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 0.1)(perc)})")), size = 0.75) +
   coord_flip() + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "County Galway",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.0),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 

```


```{r SI_map_region_of_interest_galway_city, fig.show="hold", eval = FALSE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across west of Ireland, by population"}


#Set boundings box
bbox_list <- region_of_interest %>% 
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()

pcol_si_galway_city <- si_region_of_interest_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(COUNTY_ENGLISH == "GALWAY CITY") %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/pcol_si_region_of_interest$data[1:7,2] %>% sum()) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, population, fill = transit_supply), colour="black", size=0.2) +
  scale_y_continuous(limits = c(0,500000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  population+80000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 0.1)(perc)})")), size = 0.75) +
   coord_flip() + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Galway City",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.0),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 

```





```{r SI_map_region_of_interest_county_leitrim, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across west of Ireland, by population"}


#Set boundings box
bbox_list <- region_of_interest %>% 
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()

pcol_si_county_leitrim <- si_region_of_interest_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(COUNTY_ENGLISH == "LEITRIM") %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/pcol_si_region_of_interest$data[1:7,2] %>% sum()) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, population, fill = transit_supply), colour="black", size=0.2) +
  scale_y_continuous(limits = c(0,500000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  population+80000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 0.1)(perc)})")), size = 0.75) +
   coord_flip() + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "County Leitrim (by pop.)",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.0),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 

```


```{r SI_map_region_of_interest_county_mayo, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across west of Ireland, by population"}


#Set boundings box
bbox_list <- region_of_interest %>% 
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()

pcol_si_county_mayo <- si_region_of_interest_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(COUNTY_ENGLISH == "MAYO") %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/pcol_si_region_of_interest$data[1:7,2] %>% sum()) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, population, fill = transit_supply), colour="black", size=0.2) +
  scale_y_continuous(limits = c(0,500000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  population+80000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 0.1)(perc)})")), size = 0.75) +
   coord_flip() + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "County Mayo (by pop.)",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.0),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 

```


```{r SI_map_region_of_interest_county_roscommon, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across west of Ireland, by population"}


#Set boundings box
bbox_list <- region_of_interest %>% 
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()

pcol_si_county_roscommon <- si_region_of_interest_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(COUNTY_ENGLISH == "ROSCOMMON") %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/pcol_si_region_of_interest$data[1:7,2] %>% sum()) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, population, fill = transit_supply), colour="black", size=0.2) +
  scale_y_continuous(limits = c(0,500000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  population+80000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 0.1)(perc)})")), size = 0.75) +
   coord_flip() + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "County Roscommon (by pop.)",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.0),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 

```


```{r SI_map_region_of_interest_county_sligo, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across west of Ireland, by population"}


#Set boundings box
bbox_list <- region_of_interest %>% 
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()

pcol_si_county_sligo <- si_region_of_interest_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(COUNTY_ENGLISH == "SLIGO") %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/pcol_si_region_of_interest$data[1:7,2] %>% sum()) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, population, fill = transit_supply), colour="black", size=0.2) +
  scale_y_continuous(limits = c(0,500000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  population+80000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 0.1)(perc)})")), size = 0.75) +
   coord_flip() + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "County Sligo (by pop.)",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.0),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 

```


```{r SI_map_region_of_interest_county_donegal, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across west of Ireland, by population"}


#Set boundings box
bbox_list <- region_of_interest %>% 
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()

pcol_si_county_donegal <- si_region_of_interest_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(COUNTY_ENGLISH == "DONEGAL") %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_region_of_interest_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/pcol_si_region_of_interest$data[1:7,2] %>% sum()) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, population, fill = transit_supply), colour="black", size=0.2) +
  scale_y_continuous(limits = c(0,500000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  population+80000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 0.1)(perc)})")), size = 0.75) +
   coord_flip() + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "County Donegal (by pop.)",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.0),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 

```



```{r county_boundaries, fig.show="hold", eval = FALSE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, ncol = 1, out.width="100%", fig.cap="Transit Supply Index categories across the west of Ireland, by Small Area and by population"}

county_boundaries <- region_of_interest  %>% 
    group_by(COUNTY_ENGLISH) %>%
    summarise(geometry = sf::st_union(geometry)) %>%
    ungroup()

st_write(county_boundaries, "data/west_ireland_county_boundaries.shp")

```



```{r load_county_boundaries, fig.show="hold", eval = TRUE, include = FALSE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, ncol = 1, out.width="100%", fig.cap="Transit Supply Index categories across the west of Ireland, by Small Area and by population"}

county_boundaries <- st_read("data/west_ireland_county_boundaries.shp")

```




```{r SI_map_region_of_interest_county, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, ncol = 1, out.width="100%", fig.cap="Transit Supply Index categories across the west of Ireland, by Small Area and by population"}

#Set boundings box
bbox_list <- region_of_interest %>%
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric() - 0.2*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()), bbox_list["xmax"] %>% as.numeric() )

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()


county_region_of_interest_map <- ggplot()+ 
  geom_sf(data=si_region_of_interest_aggregated$si_by_area,
          aes(fill = transit_supply), colour=NA) +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(), #remove y axis ticks
        legend.position="none",
#        legend.text = element_text(size = 4), 
#        legend.title = element_text(size = 5), 
#        legend.key.size = unit(0.08, 'cm')
        ) + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
    coord_sf(xlim = c(lon[1],lon[2]), 
           ylim = c(lat[1], lat[2])) +  
  ggspatial::annotation_north_arrow(location = 'bl', height = unit(0.1, "cm"), width = unit(0.3, "cm")) + 
  ggspatial::annotation_scale(location = 'tr', height = (unit(0.05, "cm")), text_cex = 0.25) 

county_region_of_interest_map + 
  geom_sf(data=county_boundaries, fill = NA, linewidth = 0.05) +
  inset_element(pcol_si_region_of_interest, left = 0.0, bottom = 0.8, right = 0.60, top =1.0) + 
#  inset_element(pcol_si_region_of_interest_blank, left = 0.105, bottom = 0.6, right = 0.60, top =0.8) + 
#  inset_element(pcol_si_galway_city, left = 0.113, bottom = 0.1, right = 0.608, top =0.25) + 
  inset_element(pcol_si_county_donegal, left = 0.75, bottom = 0.6, right = 1.245, top =0.75) + 
  inset_element(pcol_si_county_clare, left = 0.750, bottom = 0.0, right = 1.245, top =0.15) +  
    inset_element(pcol_si_county_galway, left = 0.750, bottom = 0.15, right = 1.245, top =0.3) +  
  inset_element(pcol_si_county_leitrim, left = 0.75, bottom = 0.45, right = 1.245, top =0.60) + 
  inset_element(pcol_si_county_mayo, left = 0.113, bottom = 0.60, right = 0.608, top =0.75) + 
  inset_element(pcol_si_county_roscommon, left = 0.750, bottom = 0.3, right = 1.245, top =0.45) + 
  inset_element(pcol_si_county_sligo, left = 0.325, bottom = 0.6, right = 0.82, top =0.75) 

``` 

In general, Figure \ref{fig:SI_map_region_of_interest_county} indicates that large parts of 
the west of Ireland have Very Low levels of transit or no transit at all. As might be expected, those Small Areas where SI scores are higher tend to be in population centres. However, of the total population across the west of Ireland (`r label_number(big.mark=",")(pcol_si_region_of_interest$data[1:7,2] %>% sum())`) `r label_percent()(pcol_si_region_of_interest$data[1:4,3] %>% sum())` (`r label_number(big.mark=",")(pcol_si_region_of_interest$data[1:4,2] %>% sum())`) live in Small Areas where the SI score is below the average (of those Small Areas with a non-zero SI score). This includes the `r label_number(big.mark=",")(pcol_si_region_of_interest$data[1,2] %>% sum())` people (`r label_percent()(pcol_si_region_of_interest$data[1,3] %>% sum())`) who live in Small Areas with no transit at all. 
The smaller bar charts in Figure \ref{fig:SI_map_region_of_interest_county} indicate that the largest shares of the total population of the west of Ireland are the `r label_number(big.mark=",")(pcol_si_county_galway$data[1,2])`, `r label_number(big.mark=",")(pcol_si_county_donegal$data[1,2])` and
`r label_number(big.mark=",")(pcol_si_county_mayo$data[1,2])` people living in County Galway, County Donegal and County Mayo in Small Areas with no transit at all.  


There is a statistically significant variation across Counties in the share of Small Areas in each transit supply category (`r prmisc::print_chi2(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% tabyl(transit_supply, COUNTY_ENGLISH) %>% chisq.test())`).
Table 3 summarises the Small Area populations in each County by transit supply category.


```{marginfigure}
Table 3: Share of populations by transit supply catergory and County.
```


```{r west_of_ireland_population_by_county, eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE}

#ggbarstats(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry(), x = transit_supply, y=COUNTY_ENGLISH, count = population)


small_area_supply_by_county <- si_region_of_interest_aggregated$si_by_area %>% 
  st_drop_geometry() %>%
  select(transit_supply, population, COUNTY_ENGLISH) %>%
  aggregate(population ~ transit_supply +  COUNTY_ENGLISH , sum)

small_area_supply_by_county_table <- small_area_supply_by_county %>% 
  pivot_wider(names_from = transit_supply, values_from = population) %>%
  mutate_all(~replace(., is.na(.), 0)) 

small_area_supply_by_county_table$total <- small_area_supply_by_county_table[,2:8] %>% rowSums()


small_area_supply_by_county_table <- small_area_supply_by_county_table[order(small_area_supply_by_county_table$total, decreasing = TRUE),]

small_area_supply_by_county_table <- small_area_supply_by_county_table [,1:8]

small_area_supply_by_county_table %>% 
  adorn_totals(where = c("row","col")) %>%
  adorn_percentages(denominator = "row") %>%
  adorn_pct_formatting() %>% 
  adorn_ns() %>% 
  select(COUNTY_ENGLISH, `Zero Supply`, `Very Low`, Low, `Below average`,`Above average`, `High`, `Very High`, Total) %>%
  kable(align = "lrrrrrrrr", 
        col.names = c("County", 
                      "Zero Supply", "Very Low", "Low",
                      "Below av.", "Above av.", 
                      "High", "Very High",
                      "Total")) %>% 
  add_header_above(c(" " = 1, "Transit Supply category" = 7, " " = 1)) %>%
  kable_styling(font_size = 8) %>%
  kableExtra::column_spec(1, width = "1.60cm") %>% 
  kableExtra::column_spec(2, width = "1.3cm") %>%
  kableExtra::column_spec(3, width = "1.3cm") %>%
  kableExtra::column_spec(4, width = "1.3cm") %>%
  kableExtra::column_spec(5, width = "1.3cm") %>%
  kableExtra::column_spec(6, width = "1.3cm")  %>%
  kableExtra::column_spec(7, width = "1.3cm") %>%
  kableExtra::column_spec(8, width = "1.3cm") %>% 
  kableExtra::column_spec(9, width = "1.3cm") 



```

Table 3 shows that  that `r label_percent(accuracy = 1)( small_area_supply_by_county_table[3,2] %>% as.numeric() / small_area_supply_by_county_table[3,2:8] %>% sum())` of people living in Mayo, `r label_percent(accuracy = 1)( small_area_supply_by_county_table[5,2] %>% as.numeric() / small_area_supply_by_county_table[5,2:8] %>% sum())` in Roscommon and `r label_percent(accuracy = 1)( small_area_supply_by_county_table[2,2] %>% as.numeric() / small_area_supply_by_county_table[2,2:8] %>% sum())` in Donegal live in Small Areas with Zero Supply, compared to just `r label_percent(accuracy = 1)( small_area_supply_by_county_table[6,2] %>% as.numeric() / small_area_supply_by_county_table[6,2:8] %>% sum())` in  Sligo, `r label_percent(accuracy = 1)( small_area_supply_by_county_table[4,2] %>% as.numeric() / small_area_supply_by_county_table[4,2:8] %>% sum())` in Clare and `r label_percent(accuracy = 1)( small_area_supply_by_county_table[1,2] %>% as.numeric() / small_area_supply_by_county_table[1,2:8] %>% sum())` in Galway. At the other end of the spectrum, some `r label_percent(accuracy = 1)( small_area_supply_by_county_table[1,6:8] %>% sum() / small_area_supply_by_county_table[1,2:8] %>% sum())` of those in Galway live in Small Areas categorised as having Above average, High or Very High transit service levels, `r label_percent(accuracy = 1)( small_area_supply_by_county_table[6,6:8] %>% sum() / small_area_supply_by_county_table[6,2:8] %>% sum())` of those in Sligo and`r label_percent(accuracy = 1)( small_area_supply_by_county_table[4,6:8] %>% sum() / small_area_supply_by_county_table[4,2:8] %>% sum())` of those in Clare. In contrast, there are no Small Areas categorised as having Above average, High or Very High transit service levels in Leitrim, while only `r label_percent(accuracy = 0.1)(small_area_supply_by_county_table[2,6:8] %>% sum() / small_area_supply_by_county_table[2,2:8] %>% sum())` of the Donegal population and `r label_percent(accuracy = 0.1)( small_area_supply_by_county_table[3,6:8] %>% sum() / small_area_supply_by_county_table[3,2:8] %>% sum())` of those in Mayo live in Small Areas with Above average, High or Very High transit service levels. This is, however, perhaps not surprising given that there are major population centres in the west of Ireland in County Galway (Galway City), County Clare (Ennis, Shannon and parts of Limerick City) and County Sligo (Sligo). However, as discussed further in the next section, Letterkenny (County Donegal), Castlebar and Ballina (County Mayo), and some other urban areas appear to have lower levels of transit supply than is the case for other places with similar (or smaller) populations.  




*Urban areas*
Approximately `r label_percent(accuracy = 10)(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME != "NA") %>% select(population) %>% sum() / si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% select(population) %>% sum())` of the west of Ireland's population (close to `r label_number(big.mark = ",", accuracy = 1000)(si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME != "NA") %>% select(population) %>% sum())` people) live in Small Areas that are within one of the `r si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME != "NA") %>% select(SA_URBAN_AREA_NAME) %>% unique() %>% nrow()` named 'urban areas' within the west. Figure \ref{fig:urban_si_pop_histogram} shows the total population and aggregate SI scores for each urban area. 

```{r urban_si_pop_histogram, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.height = 10.5, fig.width = 7.5, fig.fullwidth = TRUE, fig.cap="West of Ireland pop. vs SI, aggregated by urban areas"}



si_pop_by_urban_area <- si_region_of_interest_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME != "NA") %>% aggregate(cbind(SI, population) ~ SA_URBAN_AREA_NAME, sum)

# Some SA_URBAN_AREA_NAMES correspond to more than one COUNTY, where cities and towns have spread to both sides of a County border.  Therefore, need to create a key such that when aggregated the urban areas are allocated to the County that they are 'mostly' in.  

urban_area_to_county <- si_region_of_interest_aggregated$si_by_area %>% 
    st_drop_geometry() %>% 
    tabyl(SA_URBAN_AREA_NAME, COUNTY_ENGLISH)

urban_area_to_county <- urban_area_to_county %>%
  mutate(COUNTY_ENGLISH = case_when(
    CLARE == pmax(CLARE, DONEGAL, GALWAY, LEITRIM, MAYO, ROSCOMMON, SLIGO) ~ "CLARE",
    DONEGAL == pmax(CLARE, DONEGAL, GALWAY, LEITRIM, MAYO, ROSCOMMON, SLIGO) ~ "DONEGAL",
    GALWAY == pmax(CLARE, DONEGAL, GALWAY, LEITRIM, MAYO, ROSCOMMON, SLIGO) ~ "GALWAY",
    LEITRIM == pmax(CLARE, DONEGAL, GALWAY, LEITRIM, MAYO, ROSCOMMON, SLIGO) ~ "LEITRIM",
    MAYO == pmax(CLARE, DONEGAL, GALWAY, LEITRIM, MAYO, ROSCOMMON, SLIGO) ~ "MAYO",
    ROSCOMMON == pmax(CLARE, DONEGAL, GALWAY, LEITRIM, MAYO, ROSCOMMON, SLIGO) ~ "ROSCOMMON",
    SLIGO == pmax(CLARE, DONEGAL, GALWAY, LEITRIM, MAYO, ROSCOMMON, SLIGO) ~ "SLIGO"))

## Join to cross-ref of SA_URBAN_AREA to COUNTY_ENGLISH
si_pop_by_urban_area <- left_join(si_pop_by_urban_area, 
                                  urban_area_to_county %>% 
                                    select(SA_URBAN_AREA_NAME, COUNTY_ENGLISH))

p <- ggplot(si_pop_by_urban_area, aes(population, SI, colour = COUNTY_ENGLISH)) + 
  geom_point()  + 
  geom_text_repel(label = si_pop_by_urban_area$SA_URBAN_AREA_NAME, size = 2) +
  theme_tufte(ticks=F) + 
  scale_x_log10(labels = scales::label_number(big.mark = ",")) + 
  scale_y_log10(labels = scales::label_number(big.mark = ",")) + 
  scale_color_brewer(palette = "Dark2") +
  theme(legend.position = "bottom", 
        legend.title = element_blank()) 
  
ggMarginal(p, type = "boxplot", fill="transparent") 

#ggscatterstats(si_pop_by_urban_area, population, SI)
#gscatterstats(si_pop_by_urban_area %>% filter(SA_URBAN_AREA_NAME != "Galway city and suburbs"), population, SI)



```

There is a statistically significant relationship between population and aggregate SI score for urban areas in the west of Ireland (`r prmisc::print_cortest(cor.test(si_pop_by_urban_area$SI, si_pop_by_urban_area$population))`). However, as shown in Figure \ref{fig:urban_si_pop_histogram}, there are outliers at both ends of the spectrum of both population and aggregate SI.   



## Needs Index


```{r needs_map, fig.show="hold", eval = FALSE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="County Wexford: Small Area, by social need for transport category"}

combined_needs_2021_plot


```

```{r wexford_needs_pop_map, fig.show="hold", eval = FALSE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="County Wexford: population living in Small Areas, by social need for transport category"}

pcol_needs_wexford_county


#combined_needs_2021_plot + 
#    inset_element(pcol_needs_wexford_county, left = 0.0, bottom = 0.8, right = 0.6, top =1.0) 


```

<!---

There are `r label_number(big.mark=",")(pcol_needs_wexford_county$data[4:6,2] %>% sum())` people living in Small Areas that have social need for transport scores that are above the County Wexford average, representing some `r label_percent()(pcol_needs_wexford_county$data[4:6,3] %>% sum())` of the total popluation (`r label_number(big.mark=",")(pcol_needs_wexford_county$data[1:6,2] %>% sum())`). This includes the `r label_number(big.mark=",")(pcol_needs_wexford_county$data[6,2] %>% sum())` people (`r label_percent()(pcol_needs_wexford_county$data[6,3] %>% sum())`) living in Small Areas with Very High social needs for transport. 




\newpage


NEEDS-GAP: Table 2 shows summary statistics for SI score and transit supply category, overall and split by social needs category, for Small Areas across County Wexford.  Figure 7 shows boxplots of SI scores for Small Areas across County Wexford, by social need for transport category.  

--->

```{marginfigure}
Table 2: Small Areas: SI score and transit supply category, by needs category.
```


```{r Wexford_needs_gap_statistics, fig.show="hold", eval = FALSE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, fig.cap="County Wexford, SI score by needs category"}

si_wexford_250423_day_aggregated$si_by_area$composite_needs <- si_wexford_250423_day_aggregated$si_by_area$composite_needs %>% fct_rev()

si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>%
  select(SI, transit_supply, composite_needs) %>% 
  tbl_summary(by = composite_needs,
              digits = SI ~ 1,
              type = all_continuous() ~ "continuous2",
              statistic = list(
                all_continuous() ~ c("{min}", "{p25}", "{mean}", "{p75}", "{p100}", "{sum}"))) %>% 
  add_p(pvalue_fun = label_style_pvalue(digits = 1)) %>%
  add_overall() %>%
  # adding spanning header
  modify_spanning_header(c("stat_1", "stat_2", "stat_3", "stat_4", "stat_5", "stat_6") ~ "**Social needs category**") %>%
  as_kable_extra()



si_social_needs_test <- si_wexford_250423_day_aggregated$si_by_area %>% st_drop_geometry() %>% select(SI, composite_needs)

si_social_needs_test_result <- kruskal.test(SI ~ composite_needs, data = si_social_needs_test)

```

```{r Wexford_needs_gap_boxplot, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = FALSE, fig.cap="County Wexford, SI score by needs category"}

si_by_needs_box_plot <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>%
  select(SI, composite_needs) %>% 
  ggplot(aes(composite_needs,SI)) + 
  theme_tufte() + 
  geom_texthline(yintercept=lower_thresholds$very_high, label = "Very High", vjust=-1,color = "darkgreen", size = 2) +
  geom_texthline(yintercept=lower_thresholds$high, label = "High", vjust=-1,color = "#53A212", size = 2) +
  geom_texthline(yintercept=lower_thresholds$above_average, label = "Above Average", vjust=-0.5,color = "lightgreen", size = 2) +
  geom_texthline(yintercept=lower_thresholds$below_average, label = "Below Average", vjust=-0.5,color = "#F29C33", size = 2) +
  geom_texthline(yintercept=lower_thresholds$low, label = "Low",color = "#F7C387", size = 2) + 
  theme(axis.text.y = element_text(size = 6), 
        axis.text.x = element_text(size = 6)) +
  geom_tufteboxplot(outlier.colour="transparent") + 
  theme(axis.title=element_blank()) 

si_by_needs_box_plot


si_wexford_250423_day_aggregated$si_by_area$composite_needs <- si_wexford_250423_day_aggregated$si_by_area$composite_needs %>% fct_rev()

### Calculate by Small Area to allow statistical tests

wexford_small_area_need_gap_zone_crosstable_2021 <-  si_wexford_250423_day_aggregated$si_by_area %>% 
  st_drop_geometry() %>%
  tabyl(transit_supply,composite_needs) %>% 
  untabyl() %>% 
  select(transit_supply, `Very Low`, Low, `Below average`, `Above average`, High, `Very High`)

wexford_small_area_need_gap_zone_crosstable_2021_test <- wexford_small_area_need_gap_zone_crosstable_2021[,2:7] %>%
  chisq.test()

#wexford_small_area_need_gap_zone_crosstable_2021_test$expected
# all values over 5, assumptions met. 

```



Differences in the SI score between Small Areas in different social needs categories are statistically significant^[`r si_social_needs_test_result$method`(df=`r si_social_needs_test_result$parameter %>% as.numeric()`)=`r label_number(accuracy = 0.1)(si_social_needs_test_result$statistic %>% as.numeric())`, p=`r label_number(accuracy = 0.0001)(si_social_needs_test_result$p.value)`]. The average SI score for Small Areas with Very Low social needs for transport is `r label_number(accuracy = 0.1)(si_wexford_250423_day_aggregated$si_by_area %>% st_drop_geometry() %>% filter(composite_needs == "Very Low") %>% select(SI) %>% unlist() %>% as.numeric() %>% mean())`, compared to only
`r label_number(accuracy = 0.1)(si_wexford_250423_day_aggregated$si_by_area %>% st_drop_geometry() %>% filter(composite_needs == "Very High") %>% select(SI) %>% unlist() %>% as.numeric() %>% mean())` for those with Very High social needs for transport, which is the opposite to what might be desired to improve social equity.   


Differences in the share of Small Areas in each transit supply category across the different social need for transport groups were statistically significant^[`r prmisc::print_chi2(wexford_small_area_need_gap_zone_crosstable_2021_test)`]. However, exploring this as it relates to the population living in Small Areas with each combination of each combination of the transit supply and social need for transport categories, as shown in Table 1, may be of more interest. 





```{r Wexford_needs_gap_population, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, fig.cap="County Wexford, Populations within each SI and Combined Needs Index grouping"}

si_wexford_250423_day_aggregated$si_by_area$composite_needs <- si_wexford_250423_day_aggregated$si_by_area$composite_needs %>% fct_rev()

wexford_small_area_need_gap_population_crosstable_2021 <- si_wexford_250423_day_aggregated$si_by_area %>% 
  st_drop_geometry() %>%
  aggregate(population ~ composite_needs  + transit_supply, sum)


si_wexford_250423_day_aggregated$si_by_area$composite_needs <- si_wexford_250423_day_aggregated$si_by_area$composite_needs %>% fct_rev()


wexford_small_area_need_gap_population_crosstable_2021$composite_needs <- factor(
  wexford_small_area_need_gap_population_crosstable_2021$composite_needs, 
  levels = c("Very Low", "Low", "Below average", "Above average", "High", "Very High"))
wexford_small_area_need_gap_population_crosstable_2021$transit_supply <- factor(
  wexford_small_area_need_gap_population_crosstable_2021$transit_supply, 
  levels = c("Zero Supply", "Very Low", "Low", "Below average", "Above average", "High", "Very High"))

wexford_small_area_need_gap_population_crosstable_2021$composite_needs <- wexford_small_area_need_gap_population_crosstable_2021$composite_needs %>% 
  fct_rev()

population_supply_needs_cross_table <- wexford_small_area_need_gap_population_crosstable_2021 %>% 
  pivot_wider(names_from = composite_needs, values_from = population) 


wexford_small_area_need_gap_population_crosstable_plot <- wexford_small_area_need_gap_population_crosstable_2021 %>% 
  pivot_wider(names_from = composite_needs, values_from = population) %>%
  adorn_totals(where = c("row", "col")) %>%
  adorn_percentages(denominator = "col") %>%
  adorn_pct_formatting() %>% 
  adorn_ns() %>% 
  select(transit_supply, `Very High`, `High`, `Above average`, `Below average`, Low, `Very Low`, Total) %>%
  kable(align = "lrrrrrrr", 
        caption = "County Wexford, population in each Transit Supply and Combined Needs grouping",
        col.names = c("Transit Supply Category", 
                      "Very High", "High", 
                      "Above average", "Below average",
                      "Low", "Very Low", "Total")) %>% 
  add_header_above(c(" " = 1, "Combined Needs Index Category" = 6, " " = 1)) %>%
  kable_styling(font_size = 8) %>%
  kableExtra::column_spec(1, width = "2.5cm") %>% 
  kableExtra::column_spec(2, width = "1.25cm", background = c("#a50026", "#d73027", "#f46d43", "#fdae61", 
                                                              "#fee090","#ffffbf", "#e0f3f8","#ffffff"), color = c("white", "white", "white", "black", "black", "black", "black", "black")) %>%
  kableExtra::column_spec(3, width = "1.25cm" , background = c("#d73027", "#f46d43", "#fdae61", 
                                                               "#fee090","#ffffbf", "#e0f3f8", "#abd9e9" ,"#ffffff"), color = c("white", "white", "black", "black", "black", "black", "black", "black")) %>%
  kableExtra::column_spec(4, width = "1.25cm", background = c("#f46d43", "#fdae61", 
                                                              "#fee090","#ffffbf", "#e0f3f8", "#abd9e9","#74add1","#ffffff"), color = c("white", "black", "black", "black", "black", "black", "black", "black")) %>%
  kableExtra::column_spec(5, width = "1.25cm", background = c("#fdae61", 
                                                              "#fee090","#ffffbf", "#e0f3f8", "#abd9e9","#74add1","#4575b4", "#ffffff"), color = c("black", "black", "black", "black", "black", "black", "white", "black")) %>%
  kableExtra::column_spec(6, width = "1.25cm", background = c("#fee090","#ffffbf", "#e0f3f8", "#abd9e9","#74add1","#4575b4", "#313695", "#ffffff"), color = c("black", "black", "black", "black", "black", "white", "white", "black"))  %>%
  kableExtra::column_spec(7, width = "1.25cm", background = c("#ffffbf", "#e0f3f8", "#abd9e9","#74add1","#4575b4", "#313695", "#9400d3", "#ffffff"), color = c("black", "black", "black", "black", "white", "white", "white", "black")) %>%
  kableExtra::column_spec(8, width = "1.25cm")

wexford_small_area_need_gap_population_crosstable_2021_plot <- wexford_small_area_need_gap_population_crosstable_2021 %>%
  ggbarstats(x = transit_supply, 
             y = composite_needs, 
             counts = "population"
  ) + 
  scale_fill_manual(values = c("Very High" = "darkgreen",
                               "High" = "#53A212", 
                               "Above average" = "lightgreen", 
                               "Below average" = "#F29C33",
                               "Low" = "#F7C387", 
                               "Very Low" = "#F9D8B1", 
                               "Zero Supply" = "#FFFFFF" )) +  
  theme(legend.position="bottom")

#wexford_small_area_need_gap_population_crosstable_2021_plot


```




```{r Wexford_needs_gap_map_graphs_needs, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, fig.cap="County Wexford, Needs-gap by population, split by social needs for transport category"}

wexford_small_area_need_gap_population_crosstable_plot

```


```{r Wexford_needs_gap_map_graphs, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, fig.cap="County Wexford, Needs-gap by Small Area (red = highest gap)"}

#combine (cross) transit supply and social needs factors.
si_wexford_250423_day_aggregated$si_by_area$gap_colour <- fct_cross(si_wexford_250423_day_aggregated$si_by_area$transit_supply, si_wexford_250423_day_aggregated$si_by_area$composite_needs)

#combine colour cells
si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(b = c("Zero Supply:High", "Very Low:Very High"))

si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(c = c("Zero Supply:Above average", "Very Low:High", "Low:Very High"))

si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(d = c("Zero Supply:Below average", "Very Low:Above average", "Low:High", "Below average:Very High"))

si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(d = c("Zero Supply:Below average", "Very Low:Above average", "Low:High", "Below average:Very High"))                                                                                                                                  
si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(e = c("Zero Supply:Low", "Very Low:Below average", "Low:Above average", "Below average:High", "Above average:Very High"))                                                                                                                                  
si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(f = c("Zero Supply:Very Low", "Very Low:Low", "Low:Below average", "Below average:Above average", "Above average:High", "High:Very High"))                                            

si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(g = c("Very Low:Very Low", "Low:Low", "Below average:Below average", "Above average:Above average", "High:High", "Very High:Very High"))                                                                                                                                    
si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(h = c("Low:Very Low", "Below average:Low", "Above average:Below average", "High:Above average", "Very High:High"))                                                                                                                                                                                                                                                                      


si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(i = c("Below average:Very Low", "Above average:Low", "High:Below average", "Very High:Above average"))                                                                                                                                  
si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(j = c("Above average:Very Low", "High:Low", "Very High:Below average"))                                                                                                                               

si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_collapse(k = c("High:Very Low", "Very High:Low"))                                                                                                                                 

si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour  %>% fct_recode(a = "Zero Supply:Very High", l = "Very High:Very Low")

si_wexford_250423_day_aggregated$si_by_area$gap_colour <- si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_relevel(c("a", "b", "c", "d","e", "f", "g", "h", "i", "j", "k", "l"))




needs_gap_plot <- ggplot()+ 
  geom_sf(data=si_wexford_250423_day_aggregated$si_by_area,
          aes(fill =  gap_colour), colour=NA) +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(), #remove y axis ticks
        legend.position="none"
  ) +
  #    legend.text = element_text(size = 4), 
  #    legend.title = element_text(size = 5)) + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  ggspatial::annotation_scale(location = 'tr', height = (unit(0.05, "cm")), text_cex = 0.25) 


wexford_small_area_need_gap_population_crosstable_2021_sum_by_gap_colour <- si_wexford_250423_day_aggregated$si_by_area %>% 
  st_drop_geometry() %>%
  aggregate(population ~ gap_colour, sum)

pcol_needs_gap_wexford_county <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+5000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Needs-gap (by population)",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 2.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 


pcol_needs_gap_wexford <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Wexford") %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+4000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Wexford",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 1.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 


pcol_needs_gap_enniscorthy <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+5000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Enniscorthy",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 1.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 


pcol_needs_gap_gorey <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Gorey") %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+5000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Gorey",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 1.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 


pcol_needs_gap_new_ross <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "New Ross") %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+5000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "New Ross",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 1.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 


pcol_needs_gap_courtown <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Courtown") %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+5000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Courtown",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 1.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 


pcol_needs_gap_rosslare_harbour <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Rosslare Harbour") %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+5000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Rosslare Harbour",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 1.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 


pcol_needs_gap_castlebridge <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Castlebridge") %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+5000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Castlebridge",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 1.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 


pcol_needs_gap_rosslare <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Rosslare") %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+5000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Rosslare",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 1.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 


pcol_needs_gap_bunclody <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Bunclody") %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+5000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Bunclody",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 1.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 


pcol_needs_gap_ferns <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Ferns") %>%
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, population, fill = gap_colour), colour="black", size = 0.2) +
  scale_y_continuous(limits = c(0,60000)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  population+5000, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 0.75) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Ferns",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 1.5),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 



needs_gap_plot + 
  geom_sf(data=st_union(wexford %>% filter(SA_URBAN_AREA_NAME == "Wexford"), mapping=aes()), colour = "black", fill=NA, linewidth = 0.1) +
  geom_sf(data=st_union(wexford %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy"), mapping=aes()), colour = "black", fill=NA, linewidth = 0.1) +
  geom_sf(data=st_union(wexford %>% filter(SA_URBAN_AREA_NAME == "Gorey"), mapping=aes()), colour = "black", fill=NA, linewidth = 0.1) +
  geom_sf(data=st_union(wexford %>% filter(SA_URBAN_AREA_NAME == "Rosslare"), mapping=aes()), colour = "black", fill=NA, linewidth = 0.1) +
  geom_sf(data=st_union(wexford %>% filter(SA_URBAN_AREA_NAME == "Rosslare Harbour"), mapping=aes()), colour = "black", fill=NA, linewidth = 0.1) +
  geom_sf(data=st_union(wexford %>% filter(SA_URBAN_AREA_NAME == "New Ross"), mapping=aes()), colour = "black", fill=NA, linewidth = 0.1) +
    geom_sf(data=st_union(wexford %>% filter(SA_URBAN_AREA_NAME == "Castlebridge"), mapping=aes()), colour = "black", fill=NA, linewidth = 0.1) +
    geom_sf(data=st_union(wexford %>% filter(SA_URBAN_AREA_NAME == "Courtown"), mapping=aes()), colour = "black", fill=NA, linewidth = 0.1) +
    geom_sf(data=st_union(wexford %>% filter(SA_URBAN_AREA_NAME == "Ferns"), mapping=aes()), colour = "black", fill=NA, linewidth = 0.1) +
    geom_sf(data=st_union(wexford %>% filter(SA_URBAN_AREA_NAME == "Bunclody"), mapping=aes()), colour = "black", fill=NA, linewidth = 0.1) +
  inset_element(pcol_needs_gap_wexford_county, left = 0.0, bottom = 0.0, right = 1.0, top =0.23) + 
  inset_element(pcol_needs_gap_wexford, left = 0.70, bottom = 0.25, right = 1.70, top =0.45) + 
  inset_element(pcol_needs_gap_enniscorthy, left = 0.15, bottom = 0.58, right = 1.15, top =0.78)  +
  inset_element(pcol_needs_gap_gorey, left = 0.43, bottom = 0.82, right = 1.43, top =1.02) +
  inset_element(pcol_needs_gap_new_ross, left = 0.00, bottom = 0.46, right = 1.00, top =0.66) +
  inset_element(pcol_needs_gap_courtown, left = 0.81, bottom = 0.55, right = 1.81, top =0.75) +
  inset_element(pcol_needs_gap_rosslare_harbour, left = 0.78, bottom = 0.0, right = 1.78, top =0.20) +
    inset_element(pcol_needs_gap_castlebridge, left = 0.83, bottom = 0.37, right = 1.83, top =0.57) +
    inset_element(pcol_needs_gap_bunclody, left = 0.00, bottom = 0.8, right = 1.00, top =1.0) +
    inset_element(pcol_needs_gap_ferns, left = 0.20, bottom = 0.75, right = 1.20, top =0.95) +
  inset_element(pcol_needs_gap_rosslare, left = 0.61, bottom = 0.01, right = 1.61, top =0.21)  
  


plot_population_by_color_group_by_needs <-  function(df=df, needs_category=needs_category) {
  df %>%
    st_drop_geometry() %>%
    filter(composite_needs == needs_category) %>%
    aggregate(population ~ gap_colour, sum) %>%
    right_join(tibble(gap_colour = df$gap_colour %>% levels())) %>%
    replace_na(list(population = 0)) %>%
    mutate(gap_colour=factor(gap_colour, levels = df$gap_colour %>% fct_rev() %>% levels())) %>%
    arrange(gap_colour) %>%
    mutate(perc=population/sum(population)) %>% 
    na.omit() %>%
    ggplot() + 
    geom_col(aes(gap_colour, perc, fill = gap_colour), colour="black") +
    scale_y_continuous(limits = c(0,.6)) +
    geom_hline(yintercept = 0)  + 
    geom_text(aes(gap_colour,  0.5, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})")), size = 1.5) +
    coord_flip() + 
    scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
    scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
    guides(fill = "none", color = "none") +
    labs(
      title = needs_category,
      x = NULL,
      y = NULL
    ) +
    theme_void() +
    theme(
      panel.grid = element_blank(),
      plot.title = element_blank(),
      axis.text.y = element_blank(),
      axis.text.x = element_blank()
    ) 
}
very_high_needs_plot <- plot_population_by_color_group_by_needs(df=si_wexford_250423_day_aggregated$si_by_area, needs_category = "Very High")

high_needs_plot <- plot_population_by_color_group_by_needs(df=si_wexford_250423_day_aggregated$si_by_area, needs_category = "High")

above_average_needs_plot <- plot_population_by_color_group_by_needs(df=si_wexford_250423_day_aggregated$si_by_area, needs_category = "Above average")

below_average_needs_plot <- plot_population_by_color_group_by_needs(df=si_wexford_250423_day_aggregated$si_by_area, needs_category = "Below average")

low_needs_plot <- plot_population_by_color_group_by_needs(df=si_wexford_250423_day_aggregated$si_by_area, needs_category = "Low")

very_low_needs_plot <- plot_population_by_color_group_by_needs(df=si_wexford_250423_day_aggregated$si_by_area, needs_category = "Very Low")

```




Some `r label_number(big.mark=",")(population_supply_needs_cross_table %>% filter(transit_supply == "Zero Supply") %>% select("Very High") %>% as.numeric()) ` people in County Wexford (`r label_percent()(population_supply_needs_cross_table %>% filter(transit_supply == "Zero Supply") %>% select("Very High") %>% as.numeric() / population_supply_needs_cross_table[,-1] %>% sum())` of the total population) live in Small Areas with Very High Needs, but no transit supply. A further `r label_number(big.mark=",")(population_supply_needs_cross_table %>% filter(transit_supply == "Very Low") %>% select("Very High") %>% as.numeric())` (`r label_percent()(population_supply_needs_cross_table %>% filter(transit_supply == "Very Low") %>% select("Very High") %>% as.numeric() / population_supply_needs_cross_table[,-1] %>% sum())`) live in Small Areas with Very High Needs, but Very Low Supply, while another `r label_number(big.mark=",")(population_supply_needs_cross_table %>% filter(transit_supply == "Zero Supply") %>% select("High") %>% as.numeric())` (`r label_percent()(population_supply_needs_cross_table %>% filter(transit_supply == "Zero Supply") %>% select("High") %>% as.numeric() / population_supply_needs_cross_table[,-1] %>% sum())`) are in Small Areas with High Needs, but Zero Supply.  These people are those with the largest needs-gaps in County Wexford, but outnumber those at the other end of the spectrum^[Some `r label_number(big.mark=",")(pcol_needs_gap_wexford_county$data %>% filter(gap_colour == "l" | gap_colour == "k") %>% select(population) %>% sum())` people live in Small Areas with Very Low needs and Very High or High suppy, or Low needs and Very High supply, representing just `r label_percent()(pcol_needs_gap_wexford_county$data %>% filter(gap_colour == "l" | gap_colour == "k") %>% select(population) %>% sum() / pcol_needs_gap_wexford_county$data$population %>% sum())` of the total population.]. This relates to the general pattern that those with higher social needs for transport appear to be more likely to live in Small Areas with less transit supply. For example, 
`r label_percent()(wexford_small_area_need_gap_population_crosstable_2021[36:39,3] %>% sum() / wexford_small_area_need_gap_population_crosstable_2021[36:42,3] %>% sum())` of those living in Small Areas with Very High social needs for transport had below average transit supplies, compared to just `r label_percent()(wexford_small_area_need_gap_population_crosstable_2021[1:4,3] %>% sum() / wexford_small_area_need_gap_population_crosstable_2021[1:7,3] %>% sum())` of those living in Small Areas with Very Low social needs for transport. Out of the total `r wexford_small_area_need_gap_population_crosstable_2021[,3] %>% sum() %>% label_number(big.mark = ",")()` population of County Wexford, some `r sum(wexford_small_area_need_gap_population_crosstable_2021[36,3], wexford_small_area_need_gap_population_crosstable_2021[29,3], wexford_small_area_need_gap_population_crosstable_2021[22,3]) %>% label_number(big.mark = ",")()` (`r label_percent()(sum(wexford_small_area_need_gap_population_crosstable_2021[36,3], wexford_small_area_need_gap_population_crosstable_2021[29,3], wexford_small_area_need_gap_population_crosstable_2021[22,3]) / (wexford_small_area_need_gap_population_crosstable_2021[,3] %>% sum()))`) live in Small Areas with social needs for transport that are above the County average, but where there is no transit supply at all.  

Figure 8 maps Small Areas across County Wexford by the magnitude of the needs-gap.  In general it appears that Small Areas with larger needs gaps are in more rural areas, as might be expected given that these areas have little or no transit supply. However, at a county-wide scale it is difficult to see the detail and make comparisons between the various urban areas.  These are examined in the following. 






```{r SI_map_enniscorthy_population, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across County Wexford, by population"}


#Set boundings box
bbox_list <- wexford %>% 
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()


pcol_si_enniscorthy <- si_wexford_250423_day_aggregated$si_by_area %>%
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  st_drop_geometry() %>%
  aggregate(population ~ transit_supply, sum) %>%
  right_join(tibble(transit_supply = si_wexford_250423_day_aggregated$si_by_area$transit_supply %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(transit_supply=factor(transit_supply, levels = si_wexford_250423_day_aggregated$si_by_area$transit_supply %>% levels())) %>%
  arrange(transit_supply) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(transit_supply, perc, fill = transit_supply), colour="black") +
  scale_y_continuous(limits = c(0,.75)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(transit_supply,  0.6, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})"))) +
  coord_flip() + 
  scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Transit supply category by population",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank(),
    axis.text.y = element_text(hjust = 1),
    axis.text.x = element_blank()
  ) 



```

```{r SI_map_enniscorthy_county, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, ncol = 1, out.width="100%", fig.height = 2.5, fig.cap="Transit Supply across Enniscorthy, by Small Area"}


#Set boundings box
bbox_list <- wexford %>% 
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()


enniscorthy_map <- ggplot()+ 
  geom_sf(data=si_wexford_250423_day_aggregated$si_by_area %>% 
            filter(SA_URBAN_AREA_NAME == "Enniscorthy"),
          aes(fill = transit_supply), colour=NA) +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(), #remove y axis ticks
        legend.position="none",
        #        legend.text = element_text(size = 4), 
        #        legend.title = element_text(size = 5), 
        #        legend.key.size = unit(0.08, 'cm')
  ) + 
  scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
  coord_sf(xlim = c(lon[1],lon[2]), 
           ylim = c(lat[1], lat[2])) +  
  ggspatial::annotation_north_arrow(location = 'bl', height = unit(0.4, "cm"), width = unit(0.4, "cm")) + 
  ggspatial::annotation_scale(location = 'br', height = unit(0.1, "cm")) 


```



```{r enniscorthy_social_needs, fig.dim=c(7.5,7.5), echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE,  fig.fullwidth = FALSE,fig.cap="Wexford: Distribution of categories of social need index scores"}

needs_index_wexford <- left_join(
  wexford %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy"), 
  needs_index_wexford %>% st_drop_geometry())

combined_needs_enniscorthy_plot <- ggplot()+ 
  geom_sf(data=needs_index_wexford %>% na.omit() %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy"),
          aes(fill =  composite_needs), colour=NA) +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(), #remove y axis ticks
        legend.position="none"
  ) +
  #    legend.text = element_text(size = 4), 
  #    legend.title = element_text(size = 5)) + 
  scale_fill_manual(values = c("#FFFB07", "#CEE102", "#A3CF07", "#61AE09", "#2EA105", "#0A7B01")) +
  ggspatial::annotation_scale(location = 'br') 

#combined_needs_2021_plot

```



```{r enniscorthy_social_needs_population, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Transit Supply across County Wexford, by population"}


#Set boundings box
bbox_list <- wexford  %>% 
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  st_geometry %>% 
  st_combine() %>%
  st_bbox()


lat = c(bbox_list["ymin"] %>% as.numeric(), bbox_list["ymax"] %>% as.numeric())
lon = c(bbox_list["xmin"] %>% as.numeric(), bbox_list["xmax"] %>% as.numeric() + 0*(bbox_list["xmax"] %>% as.numeric() - bbox_list["xmin"] %>% as.numeric()))

Poly_Coord_df = data.frame(lon, lat)

poly <- Poly_Coord_df %>% 
  st_as_sf(coords = c("lon", "lat"), 
           crs = 2157) %>% 
  st_bbox() %>% 
  st_as_sfc()


pcol_needs_enniscorthy <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>%
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  aggregate(population ~ composite_needs, sum) %>%
  right_join(tibble(composite_needs = si_wexford_250423_day_aggregated$si_by_area$composite_needs %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(composite_needs=factor(composite_needs, levels = si_wexford_250423_day_aggregated$si_by_area$composite_needs %>% levels())) %>%
  arrange(composite_needs) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(composite_needs, perc, fill = composite_needs), colour="black") +
  scale_y_continuous(limits = c(0,.6)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(composite_needs,  0.5, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})"))) +
  coord_flip() + 
  scale_fill_manual(values = c("Very Low" = "#FFFB07", "Low" = "#CEE102", "Below average" = "#A3CF07", "Above average" = "#61AE09", "High" = "#2EA105", "Very High" = "#0A7B01")) + 
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Transport need by population",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank(),
    axis.text.y = element_text(hjust = 1),
    axis.text.x = element_blank()
  ) 



```




TRANSIT SUPPLY:  Figure 9 maps Small Areas across Enniscorthy by transit supply category.  Table 3 shows summary statistics for SI score and population, overall and split by transit supply category. 



```{marginfigure}
Table 3: Enniscorthy: summary statistics, SI \& pop.
```


```{r enniscorthy _si_table_summary, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = FALSE, fig.cap="Enniscorthy: population living in Small Areas, by transit supply category"}

theme_gtsummary_compact()
si_wexford_250423_day_aggregated$si_by_area %>% 
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  select(SI, population, transit_supply) %>% 
  droplevels() %>%   
  tbl_summary(
    by = transit_supply,
    digits = SI ~ 1,
    type = all_continuous() ~ "continuous2",
    statistic = list(
      all_continuous() ~ c("{min}", "{p25}", "{mean}", "{p75}", "{p100}", "{sum}")
    )) %>%
  add_overall() %>% 
  modify_spanning_header(c("stat_1", "stat_2", "stat_3", "stat_4", "stat_5") ~ "**Transit supply category**") %>%
  as_kable_extra() %>%
  kableExtra::column_spec(2, width = "1.0cm") %>%
  kableExtra::column_spec(3, width = "1.25cm") %>%  
  kableExtra::column_spec(4, width = "1.25cm") %>%
  kableExtra::column_spec(5, width = "1.25cm") %>%
  kableExtra::column_spec(6, width = "1.0cm") %>%  
  kableExtra::column_spec(7, width = "1.25cm") 




text <- si_wexford_250423_day_aggregated$thresholds[2]
numbers <- regmatches(text,gregexpr("[-+]?[0-9]*\\.?[0-9]+([eE][-+]?[0-9]+)?",text)) %>% 
  unlist() %>% 
  as.numeric()
minimum_value_of_threshold <- numbers[1]


```



```{r enniscorthy_si_map, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Enniscorthy: Small Areas by transit supply category"}
enniscorthy_map 

```

```{r enniscorthy_si_population, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Enniscorthy: population by transit supply category"}
pcol_si_enniscorthy  

```


The minimum SI score for Small Areas within Enniscorthy is `r label_number(accuracy = 0.01)(si_wexford_250423_day_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>% select(SI) %>% min())`, which is just higher than the upper threshold of the "Very Low" transit supply category (`r label_number(accuracy = 0.01)(minimum_value_of_threshold)`). Hence, none of the `r label_number(big.mark=",")(pcol_si_enniscorthy$data[1:7,2] %>% sum())` Enniscorthy residents live in Small Areas with "Very Low" or "Zero Supply" transit supply, although `r label_percent()(pcol_si_enniscorthy$data[1:4,3] %>% sum())` (`r label_number(big.mark=",")(pcol_si_enniscorthy$data[1:4,2] %>% sum())`) live in Small Areas where the SI score is below the County Wexford average (`r label_number(accuracy = 0.1)(si_wexford_250423_day_aggregated$si_by_area$SI %>% mean())`). 

```{r enniscorthy_gglorenz, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Enniscorthy: Lorenz curve"}




perc_population <- pcol_si_enniscorthy$data %>% 
  select(transit_supply, population) %>%
  adorn_percentages(denominator = "col") 

perc_population$cumulative_pop <- perc_population$population %>% cumsum()


si_wexford_250423_day_aggregated$si_by_area %>%
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  ggplot(aes(x = SI, n = population)) +
  stat_lorenz(desc = FALSE) +
  coord_fixed() +
  geom_abline(linetype = "dashed") +
  theme_minimal() +
  hrbrthemes::scale_x_percent() +
  hrbrthemes::scale_y_percent() +
  labs(x = "Cumulative share of population ",
       y = "Cumulative share of SI score") + 
  geom_textvline(xintercept = perc_population[6,3], vjust=0.5, color = "darkgreen", label = "Very High") +
  geom_textvline(xintercept = perc_population[5,3], vjust=0.5, color = "#53A212", label = "High") +
  geom_textvline(xintercept = perc_population[4,3], vjust=1.5, color = "lightgreen", label = "Above Average") +
  geom_textvline(xintercept = perc_population[3,3], vjust=2.5, color = "#F29C33", label = "Below Average") +
  geom_textvline(xintercept = perc_population[2,3], vjust=1.7, color = "#F7C387", label = "Low") +
  geom_textvline(xintercept = perc_population[1,3], vjust=1.5, color = "#F9D8B1", label = "")

lorenz_values <- si_wexford_250423_day_aggregated$si_by_area %>% 
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  select(SI, population)
lorenz_values <- Lc(lorenz_values$SI, lorenz_values$population)

```


A Lorenz Curve of distribution of Enniscorthy's transit supply to the population is shown in Figure 11. The Gini co-efficient is `r lorenz_values$Gini %>% label_number(accuracy=0.01)()`, indicating that transit supply is more evenly distributed to the population than is the case across County Wexford as a whole. The SI score aggregated all of Enniscorthy is approximately `r label_number(big.mark=",", accuracy = 100)(si_wexford_250423_day_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>% select(SI) %>% sum())`, but `r label_percent(digits=2)(1-lorenz_values$L[40])` of this goes to Small Areas that are home to just `r label_percent(digits=2)(1-lorenz_values$p[40])` of the population. `r label_percent(digits=2)(lorenz_values$p[8])` of the population live in Small Areas that, in combination, receive just `r label_percent(digits=2)(lorenz_values$L[8])` of the total supply. 


```{r enniscorthy_si_pop_histogram, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = FALSE, fig.cap="Enniscorthy: SI and distance (from Small Area's centroid) to post office, sized by population"}





df <- data.frame(
  lat = c(52.50104027729214),
  lon = c(-6.566539225188304))

enniscorthy_main_post_office <- df %>% st_as_sf(coords = c("lon", "lat"), crs = 4326)

enniscorthy_main_post_office  <- st_transform(enniscorthy_main_post_office, 2157)


enniscorthy_only <- si_wexford_250423_day_aggregated$si_by_area %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy")


enniscorthy_only$distance_m_to_GPO <- enniscorthy_only %>% st_centroid() %>% 
  st_distance(enniscorthy_main_post_office) %>% 
  as.numeric()




q <- ggplot(enniscorthy_only %>% st_drop_geometry(), aes(distance_m_to_GPO, SI)) + 
  geom_point(aes(size=population, fill=transit_supply), show.legend = "none", shape=21)  + 
scale_fill_manual(values = c("Zero Supply" = "white", "Very Low" = "#F9D8B1", "Low" = "#F7C387", "Below average" = "#F29C33", "Above average" = "lightgreen", "High" = "#53A212", "Very High" = "darkgreen")) +
theme_tufte(ticks=F) + 
  scale_x_log10(labels = scales::label_number(big.mark = ",")) + 
  scale_y_log10(labels = scales::label_number(big.mark = ",")) 
ggMarginal(q, type = "histogram", fill="transparent") 



```

In general, and as shown in Figure 9, the transit supply in Enniscorthy appears to be higher in more central areas. Figure 12 plots SI score against distance to the main post office in Enniscorthy, which indicates that transit supply is High or Very High for almost all Small Areas with centroids closer than around 500 metres, yet these areas generally appear to have small populations.  Further away from the centre of town, however, there appears to be less of a link between distance and transit supply. Notably there appears to be a swathe of Small Areas with "Above Average" transit supply and  populations in the 200-250 range, mostly located along a WNW corridor^[Bellefield Road (R702).], whereas northern, . 

SOCIAL NEEDS FOR TRANSPORT: The average social needs index score across Enniscorthy is `r label_number(accuracy = 0.1)(enniscorthy_only$combined_needs_index %>% mean())`. Figure 13 maps the social needs categories across County Wexford, while Figure 14 shows the grouping by population. 


```{r enniscorthy_needs_map, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Enniscorthy: Small Area, by social need for transport category"}

combined_needs_enniscorthy_plot


```

```{r enniscorthy_needs_pop_map, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Enniscorthy: population living in Small Areas, by social need for transport category"}

pcol_needs_enniscorthy


#combined_needs_2021_plot + 
#    inset_element(pcol_needs_wexford_county, left = 0.0, bottom = 0.8, right = 0.6, top =1.0) 


```


There are `r label_number(big.mark=",")(pcol_needs_enniscorthy$data[4:6,2] %>% sum())` people living in Small Areas that have social need for transport scores that are above the County Wexford average, representing some `r label_percent()(pcol_needs_enniscorthy$data[4:6,3] %>% sum())` of the Enniscorthy popluation (`r label_number(big.mark=",")(pcol_needs_enniscorthy$data[1:6,2] %>% sum())`). This includes the `r label_number(big.mark=",")(pcol_needs_enniscorthy$data[6,2] %>% sum())` people (`r label_percent()(pcol_needs_enniscorthy$data[6,3] %>% sum())`) living in Small Areas with Very High social needs for transport. 



\newpage


NEEDS-GAP: Table 4 shows summary statistics for SI score and transit supply category, overall and split by social needs category, for Small Areas across Enniscorthy.  Figure 15 shows boxplots of SI scores for Small Areas across Enniscorthy, by social need for transport category.  

```{marginfigure}
Table 4: Enniscorthy Small Areas: SI score and transit supply category, by needs category.
```


```{r Enniscorthy_needs_gap_statistics, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, fig.cap="County Wexford, SI score by needs category"}

si_wexford_250423_day_aggregated$si_by_area$composite_needs <- si_wexford_250423_day_aggregated$si_by_area$composite_needs %>% fct_rev()

si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  select(SI, transit_supply, composite_needs) %>% 
  droplevels() %>%  
  tbl_summary(by = composite_needs,
              digits = SI ~ 1,
              type = all_continuous() ~ "continuous2",
              statistic = list(
                all_continuous() ~ c("{min}", "{p25}", "{mean}", "{p75}", "{p100}", "{sum}"))) %>% 
  add_p(pvalue_fun = label_style_pvalue(digits = 1)) %>%
  add_overall() %>%
  # adding spanning header
  modify_spanning_header(c("stat_1", "stat_2", "stat_3", "stat_4", "stat_5", "stat_6") ~ "**Social needs category**") %>%
  as_kable_extra() 


si_social_needs_test <- si_wexford_250423_day_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>% select(SI, composite_needs)

si_social_needs_test_result <- kruskal.test(SI ~ composite_needs, data = si_social_needs_test)

```

```{r Enniscorthy_needs_gap_boxplot, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = FALSE, fig.cap="Enniscorthy, SI score by needs category"}

si_by_needs_box_plot <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  select(SI, composite_needs) %>% 
  ggplot(aes(composite_needs,SI)) + 
  theme_tufte() + 
  geom_texthline(yintercept=lower_thresholds$very_high, label = "Very High", vjust=-1,color = "darkgreen", size = 2) +
  geom_texthline(yintercept=lower_thresholds$high, label = "High", vjust=-1,color = "#53A212", size = 2) +
  geom_texthline(yintercept=lower_thresholds$above_average, label = "Above Average", vjust=-0.5,color = "lightgreen", size = 2) +
  geom_texthline(yintercept=lower_thresholds$below_average, label = "Below Average", vjust=-0.5,color = "#F29C33", size = 2) +
  geom_texthline(yintercept=lower_thresholds$low, label = "Low",color = "#F7C387", size = 2) + 
  theme(axis.text.y = element_text(size = 6), 
        axis.text.x = element_text(size = 6)) +
  geom_tufteboxplot(outlier.colour="transparent") + 
  theme(axis.title=element_blank()) 

si_by_needs_box_plot


si_wexford_250423_day_aggregated$si_by_area$composite_needs <- si_wexford_250423_day_aggregated$si_by_area$composite_needs %>% fct_rev()

### Calculate by Small Area to allow statistical tests

enniscorthy_small_area_need_gap_zone_crosstable_2021 <-  si_wexford_250423_day_aggregated$si_by_area %>% 
  st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  tabyl(transit_supply,composite_needs) %>% 
  untabyl() %>% 
  select(transit_supply, `Very Low`, Low, `Below average`, `Above average`, High, `Very High`)

enniscorthy_small_area_need_gap_zone_crosstable_2021_test <- enniscorthy_small_area_need_gap_zone_crosstable_2021[3:7,2:7] %>%
  fisher.test(simulate.p.value=TRUE)

enniscorthy_small_area_need_gap_zone_crosstable_2021_test$expected

```



Differences in the SI score between Small Areas in different social needs categories are statistically significant^[`r si_social_needs_test_result$method`(df=`r si_social_needs_test_result$parameter %>% as.numeric()`)=`r label_number(accuracy = 0.1)(si_social_needs_test_result$statistic %>% as.numeric())`, p=`r label_number(accuracy = 0.0001)(si_social_needs_test_result$p.value)`]. The average SI score for Small Areas with Very Low social needs for transport is `r label_number(accuracy = 0.1)(si_wexford_250423_day_aggregated$si_by_area %>% st_drop_geometry() %>% filter(composite_needs == "Very Low") %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>% select(SI) %>% unlist() %>% as.numeric() %>% mean())`, compared to only
`r label_number(accuracy = 0.1)(si_wexford_250423_day_aggregated$si_by_area %>% st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%  filter(composite_needs == "Very High") %>% select(SI) %>% unlist() %>% as.numeric() %>% mean())` for those with Very High social needs for transport, which is the opposite to what might be desired to improve social equity.   


Differences in the share of Small Areas in each transit supply category across the different social need for transport groups were statistically significant^[`r enniscorthy_small_area_need_gap_zone_crosstable_2021_test$method`, p=`r label_number(accuracy = 0.001)(enniscorthy_small_area_need_gap_zone_crosstable_2021_test$p.value)`]. Again, this is explored further with respect to the aggregate population in each combination of categories in the following.




```{r Enniscorthy_needs_gap_population, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, fig.cap="County Wexford, Populations within each SI and Combined Needs Index grouping"}

si_wexford_250423_day_aggregated$si_by_area$composite_needs <- si_wexford_250423_day_aggregated$si_by_area$composite_needs %>% fct_rev()

enniscorthy_small_area_need_gap_population_crosstable_2021 <- 
  xtabs(population ~ transit_supply + composite_needs, si_wexford_250423_day_aggregated$si_by_area %>% 
          st_drop_geometry() %>% 
          filter(SA_URBAN_AREA_NAME == "Enniscorthy")) %>%
  as.tibble()

names(enniscorthy_small_area_need_gap_population_crosstable_2021) <- c("transit_supply", "composite_needs", "population")


population_supply_needs_cross_table <- enniscorthy_small_area_need_gap_population_crosstable_2021 %>% 
  pivot_wider(names_from = composite_needs, values_from = population) 


enniscorthy_small_area_need_gap_population_crosstable_plot <- enniscorthy_small_area_need_gap_population_crosstable_2021 %>% 
  pivot_wider(names_from = composite_needs, values_from = population, names_expand = TRUE) %>%
  adorn_totals(where = c("row", "col")) %>%
  adorn_percentages(denominator = "col") %>%
  adorn_pct_formatting() %>% 
  adorn_ns() %>% 
  select(transit_supply, `Very High`, `High`, `Above average`, `Below average`, Low, `Very Low`, Total) %>%
  kable(align = "lrrrrrrr", 
        caption = "Enniscorthy, population in each Transit Supply and Combined Needs grouping",
        col.names = c("Transit Supply Category", 
                      "Very High", "High", 
                      "Above average", "Below average",
                      "Low", "Very Low", "Total")) %>% 
  add_header_above(c(" " = 1, "Combined Needs Index Category" = 6, " " = 1)) %>%
  kable_styling(font_size = 8) %>%
  kableExtra::column_spec(1, width = "2.25cm") %>% 
  kableExtra::column_spec(2, width = "1.0cm", background = c("#a50026", "#d73027", "#f46d43", "#fdae61", 
                                                             "#fee090","#ffffbf", "#e0f3f8","#ffffff"), color = c("white", "white", "white", "black", "black", "black", "black", "black")) %>%
  kableExtra::column_spec(3, width = "1.0cm" , background = c("#d73027", "#f46d43", "#fdae61", 
                                                              "#fee090","#ffffbf", "#e0f3f8", "#abd9e9" ,"#ffffff"), color = c("white", "white", "black", "black", "black", "black", "black", "black")) %>%
  kableExtra::column_spec(4, width = "1.0cm", background = c("#f46d43", "#fdae61", 
                                                             "#fee090","#ffffbf", "#e0f3f8", "#abd9e9","#74add1","#ffffff"), color = c("white", "black", "black", "black", "black", "black", "black", "black")) %>%
  kableExtra::column_spec(5, width = "1.0cm", background = c("#fdae61", 
                                                             "#fee090","#ffffbf", "#e0f3f8", "#abd9e9","#74add1","#4575b4", "#ffffff"), color = c("black", "black", "black", "black", "black", "black", "white", "black")) %>%
  kableExtra::column_spec(6, width = "1.0cm", background = c("#fee090","#ffffbf", "#e0f3f8", "#abd9e9","#74add1","#4575b4", "#313695", "#ffffff"), color = c("black", "black", "black", "black", "black", "white", "white", "black"))  %>%
  kableExtra::column_spec(7, width = "1.0cm", background = c("#ffffbf", "#e0f3f8", "#abd9e9","#74add1","#4575b4", "#313695", "#9400d3", "#ffffff"), color = c("black", "black", "black", "black", "white", "white", "white", "black")) %>%
  kableExtra::column_spec(8, width = "1.0cm")

enniscorthy_small_area_need_gap_population_crosstable_2021_plot <- enniscorthy_small_area_need_gap_population_crosstable_2021 %>%
  ggbarstats(x = transit_supply, 
             y = composite_needs, 
             counts = "population"
  ) + 
  scale_fill_manual(values = c("Very High" = "darkgreen",
                               "High" = "#53A212", 
                               "Above average" = "lightgreen", 
                               "Below average" = "#F29C33",
                               "Low" = "#F7C387", 
                               "Very Low" = "#F9D8B1", 
                               "Zero Supply" = "#FFFFFF" )) +  
  theme(legend.position="bottom")


```




```{r Enniscorthy_needs_gap_map_graphs_needs, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.fullwidth = TRUE, fig.cap="County Wexford, Needs-gap by population, split by social needs for transport category"}

enniscorthy_small_area_need_gap_population_crosstable_plot



pcol_needs_gap_enniscorthy <- si_wexford_250423_day_aggregated$si_by_area %>%
  st_drop_geometry() %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>% 
  aggregate(population ~ gap_colour, sum) %>%
  right_join(tibble(gap_colour = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% levels())) %>%
  replace_na(list(population = 0)) %>%
  mutate(gap_colour=factor(gap_colour, levels = si_wexford_250423_day_aggregated$si_by_area$gap_colour %>% fct_rev() %>% levels())) %>%
  arrange(gap_colour) %>%
  mutate(perc=population/sum(population)) %>% 
  na.omit() %>%
  ggplot() + 
  geom_col(aes(gap_colour, perc, fill = gap_colour), colour="black") +
  scale_y_continuous(limits = c(0,.5)) +
  geom_hline(yintercept = 0)  + 
  geom_text(aes(gap_colour,  0.4, label = glue::glue("{label_comma(accuracy = 1, big.mark = ",")(population)} ({label_percent(accuracy = 1)(perc)})"))) +
  coord_flip() + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  scale_color_manual(values = c(rep("black", 5), rep("white", 2), "black")) +
  guides(fill = "none", color = "none") +
  labs(
    title = "Needs-gap by population",
    x = NULL,
    y = NULL
  ) +
  theme_void() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank()
  ) 

#needs_gap_plot + 
#    inset_element(pcol_needs_gap_wexford_county, left = 0.0, bottom = 0.8, right = 0.6, top =1.0) 



```

Figure 16 summarises the magnitude of the needs-gap across the Enniscorthy population, while Figure 17 maps the needs-gaps. Although there are no Small Areas in Enniscorthy in the two largest needs-gaps categories (dark red, red), There are `r label_number(big.mark = ",")(pcol_needs_gap_enniscorthy$data[7:12,2] %>% sum())` people living in the Small Areas in the upper half of the needs-gaps categories (i.e. yellow or orange), being `r label_percent(accuracy = 1)(pcol_needs_gap_enniscorthy$data[7:12,2] %>% sum() / pcol_needs_gap_enniscorthy$data[,2] %>% sum())` of the total Enniscorthy population. As shown in the map in Figure 17, these people appear to mostly living in outer parts of Enniscorthy.  


```{r enniscorthy_needs_gap_map_poplation, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Enniscorthy: Needs-gap by population"}
pcol_needs_gap_enniscorthy

```


```{r enniscorthy_needs_gap_map_graphs, fig.show="hold", eval = TRUE, echo = FALSE, warning=FALSE, message=FALSE, cache=FALSE, fig.margin = TRUE, fig.cap="Enniscorthy: Needs-gap by Small Area"}


needs_gap_plot <- ggplot()+ 
  geom_sf(data=si_wexford_250423_day_aggregated$si_by_area %>% filter(SA_URBAN_AREA_NAME == "Enniscorthy"),
          aes(fill =  gap_colour), colour=NA) +
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank(), #remove y axis ticks
        legend.position="none"
  ) +
  #    legend.text = element_text(size = 4), 
  #    legend.title = element_text(size = 5)) + 
  scale_fill_manual(values = c(a = "#a50026", b = "#d73027", c ="#f46d43", d="#fdae61", e="#fee090", f="#ffffbf", g="#e0f3f8", h="#abd9e9", i= "#74add1", j="#4575b4", k="#313695", l="#9400d3")) +
  ggspatial::annotation_scale(location = 'br') 


enniscorthy_small_area_need_gap_population_crosstable_2021_sum_by_gap_colour <- si_wexford_250423_day_aggregated$si_by_area %>% 
  st_drop_geometry() %>% 
  filter(SA_URBAN_AREA_NAME == "Enniscorthy") %>%
  aggregate(population ~ gap_colour, sum)

needs_gap_plot  
```




```{r, include=FALSE}
knitr::write_bib(file = 'packages.bib')
```
 



